---
title: "MVLN "
subtitle: "**North Atlantic Blue Shark**"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
# bibliography: refs.bib
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../outputs/html") })
---
  
```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# getwd()
library(knitr)
opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               cache     =TRUE,
               cache.path="cache/mvln/",
               fig.path  ="../outputs/tex/mvln-",
               fig.width =10,
               fig.height=6,
               dev       ="png")
iFig=0
iTab=0
```


```{r}
pagebreak <- function() {
  if(knitr::is_latex_output())
    return("\\newpage")
  else
    return('<div style="page-break-before: always;" />')
}
```


```{r, warning=FALSE,  message=FALSE, error=FALSE}
library(r4ss)
library(ss3diags)     

library(FLCore)
library(ggplotFL)
library(kobe)

library(ggplot2)
library(grid) 

library(reshape)
library(plyr)
library(dplyr)

library(data.table)

library(xtable)
library(kableExtra)
```

```{r, dirMy, eval=FALSE}
setwd("~/Desktop/SCRS-papers/mvln/Rmd")
```

```{r, source, echo=FALSE}
source("../R/ssMVLNFunction.R")  
```

```{r, base, echo=FALSE}
base=SS_output("../data/inputs/run_06",verbose=FALSE,hidewarn=FALSE,printstat=FALSE)
```

```{r, mvln}
mvln =mvlnFLQ(ssmvln(base[["CoVar"]],base[["derived_quants"]],mc= 1241))
```

## Historical Status

```{r, trend-stock, fig.width=8, fig.height=4}
kobe:::trendPhaseMar2(window(mvln[["stock"]],end=2022),ylim=3, ylab=expression(SSB:B[MSY]))
```

**Figure `r iFig=iFig+1; iFig`.** Historical trends in $SSB/B_{MSY}$ with marginal plot of current status, ribbons show inter-quartiles and $90_{th}$ percentile, relative to median. 


```{r, trend-harvest, fig.width=8, fig.height=4, echo=FALSE}
kobe:::trendPhaseMar2(window(mvln[["harvest"]],end=2022),pts=mvln[["harvest"]][,"2023"],ylim=2,
               ylab=expression(F:F[MSY]))   
```

**Figure `r iFig=iFig+1; iFig`.** Historical trends in $F/F_{MSY}$ with marginal plot of current status, ribbons show inter-quartiles and $90_{th}$ percentile, relative to median. 

## Current Status

```{r, fig.height=8, fig.width=8}
kb=rbind(cbind(run="mvln",model.frame(FLQuants(stock=mvln[["stock"]][,"2021"],
                                               harvest=mvln[["harvest"]][,"2020"]),drop=T)))

kobe:::kobePhaseMar2(kb,
                     quadcol=c("red","green","yellow","orange"),
                     col="blue",size=0.1) 
```

**Figure `r iFig=iFig+1; iFig`.** Kobe phase plot for $SSB/B_{MSY}$ in 2021 and $F/F_{MSY}$ in 2020

```{r, fig.height=8, fig.width=8} 
kb=rbind(cbind(run="mvln",model.frame(FLQuants(stock=mvln[["stock"]][,"2022"],
                                               harvest=mvln[["harvest"]][,"2022"]),drop=T)))

kobe:::kobePhaseMar2(kb, quadcol=c("red","green","yellow","orange"),
                     col="blue",size=0.1) 
```

**Figure `r iFig=iFig+1; iFig`.** Kobe phase plot for $SSB/B_{MSY}$ in 2022 and $F/F_{MSY}$ in 2022


```{r, kobes-mvln,fig.height=8, fig.width=8}
dat=mdply(expand.grid(F=2020:2023,SSB=2020:2023), function(F,SSB) 
  model.frame(FLQuants(harvest=mvln[["harvest"]][,ac(F)],
                       stock  =mvln[["stock"]][  ,ac(SSB)]),drop=T))
kobe::kobePhase(dat)+
  geom_point(aes(stock,harvest),shape=21,fill="grey",col="black",size=0.75,linewidth=0.3)+
  geom_label(aes(x=0.2,y=0.1,label=signif(V1,2)),
             data=ddply(dat,.(F,SSB), with, cor(log(stock),log(harvest))))+
  facet_grid(F~SSB)
```

**Figure `r iFig=iFig+1; iFig`.** Kobe phase plots for historical and $F/F_{MSY}$ forecasts, with correlations. 

## Projections


```{r, bsh300-mvln}
dir300="../data/inputs/forecasts/300"

bsh300=SS_output(dir300,verbose=FALSE,hidewarn=FALSE,printstat=FALSE)
bsh300=bsh300[c("CoVar","derived_quants")]

mvln =mvlnFLQ(ssmvln(bsh300[["CoVar"]],bsh300[["derived_quants"]],mc= 1241)) 
```

```{r, trend-stock-hist, fig.width=8, fig.height=4,warning=FALSE, message=FALSE}
kobe:::trendPhaseMar2(window(mvln[["stock"]],start=2000, end=2070),ylim=3, 
               ylab=expression(SSB:B[MSY]))
```

**Figure `r iFig=iFig+1; iFig`.** Recent historical and future trends in $SSB/B_{MSY}$ and projection for $F_{MSY}$, marginal plot of status in 2032, ribbons show inter-quartiles and $90_{th}$ percentile, relative to median.


```{r, trend-harvest-hist, fig.width=8, fig.height=4, echo=FALSE}
kobe:::trendPhaseMar2(window(mvln[["harvest"]],start=2000, end=2070),ylim=2, 
               ylab=expression(F:F[MSY]))
```

**Figure `r iFig=iFig+1; iFig`.** Recent historical and future trends in $F/F_{MSY}$ and projection for $F_{MSY}$, marginal plot of status in 2032, ribbons show inter-quartiles and $90_{th}$ percentile, relative to median.


```{r, cvrs}
dir="../data/inputs/forecasts"

runs=c(0,seq(150,400,25))

cvrs=mlply(runs, function(x) try(SS_output(file.path(dir,x),verbose=FALSE,
                                           hidewarn=FALSE,printstat=FALSE)$CoVar)) 
hats=mlply(runs, function(x) try(SS_output(file.path(dir,x),verbose=FALSE,
                                           hidewarn=FALSE,printstat=FALSE)$derived_quants))
```

```{r, ts}
hat   =transform(ldply(hats, function(x) if(is.data.frame(x)) return(x) else return(NULL)) ,Run=runs[an(ac(X1))])
bratio=transform(hat[grep("Bratio_",hat$Label),],year=an(substr(Label,8,12)))
ssb   =subset(transform(hat[grep("SSB_",   hat$Label),],year=an(substr(Label,5,nchar(Label)))),!is.na(year))
f     =subset(transform(hat[grep("F_",     hat$Label),],year=an(substr(Label,3,7))),           !is.na(year))
af    =transform(hat[grep("annF_",  hat$Label),],ref=an(substr(Label,6,nchar(Label))))
```

```{r, bratio, fig.width=8,fig.height=4}
ggplot(transform(bratio,What=ifelse(substr(Run,1,1)=="F","F","Catch"),TAC=paste(Run,"(kt)")))+
  scale_x_continuous(limits=c(1970,2072))+
  theme_bw()+
  theme(legend.position="bottom")+
  geom_line(aes(year,Value,col=TAC))+
  xlab("Year")+ylab(expression(SSB/B[MSY]))+
  scale_color_manual("TAC",values=rainbow(length(unique(f$Run))))
```

**Figure `r iFig=iFig+1; iFig`.** Deterministic values of $SSB/B_{MSY}$ for projections

```{r, f, fig.width=8,fig.height=4}
ggplot(transform(f,What=ifelse(substr(Run,1,1)=="F","F","Catch"),TAC=paste(Run,"(kt)")))+  
  scale_y_continuous(limits=c(0,3))+
  scale_x_continuous(limits=c(1970,2072))+
  geom_line(aes(year,Value,col=TAC))+
  scale_y_continuous(limits=c(0,2))+
  scale_x_continuous(limits=c(1970,2072))+
  theme_bw()+
  theme(legend.position="bottom")+
  xlab("Year")+ylab(expression(F/F[MSY]))+
  scale_color_manual("TAC",values=rainbow(length(unique(f$Run))))
```

**Figure `r iFig=iFig+1; iFig`.** Deterministic values of $F/F_{MSY}$ for projections.

```{r, mvls, fig.width=8,fig.height=4}
mvls=mlply(seq(length(cvrs)), function(x) 
         {set.seed(1224); try(model.frame(mvlnFLQ(ssmvln(cvrs[[x]],hats[[x]],mc= 1241)),drop=T))}) 
dat=transform(ldply(mvls, function(x) if(is.data.frame(x)) return(x) else return(NULL)),Run=runs[an(ac(X1))])
dt2=transform(ddply(dat, .(year,Run), with, median(stock)),What=ifelse(substr(Run,1,1)=="F","F","Catch"))

ggplot(transform(dt2,TAC=paste(Run,"(kt)")))+
  geom_line(aes(x=year,y=V1,col=TAC))+
  scale_x_continuous(limits=c(1970,2072))+
  theme_bw()+
  theme(legend.position="bottom")+
  xlab("Year")+ylab(expression(SSB/B[MSY]))+
  scale_color_manual("TAC",values=rainbow(length(unique(f$Run))))
```

**Figure `r iFig=iFig+1; iFig`.**  Median estimates of $B/SSB_{MSY}$ for projections. 

```{r, mvls-f, fig.width=8,fig.height=4}
dt2=transform(ddply(dat, .(year,Run), with,    
                    median(harvest)),What=ifelse(substr(Run,1,1)=="F","F","Catch"))

ggplot(transform(dt2,TAC=paste(Run,"(kt)")))+
  geom_line(aes(x=year,y=V1,col=TAC))+
  xlab("Year")+ylab(expression(F/F[MSY]))+
  scale_y_continuous(limits=c(0,3))+
  scale_x_continuous(limits=c(1970,2072))+
  facet_grid(What~.)+
  theme_bw()+
  theme(legend.position="bottom")+
  scale_color_manual("TAC",values=rainbow(length(unique(f$Run))))
```

**Figure `r iFig=iFig+1; iFig`.**  Median estimates of $F/SSB_{MSY}$ for projections. 

```{r}
library(ggplotFL)
plot(mvln,iter=c(5,17))+
  theme_bw()
```

**Figure `r iFig=iFig+1; iFig`.**  Plot of $B/SSB_{MSY}$ & $F/F_{MSY}$ with sample Monte Carlo simulations.

`r pagebreak()`

\tiny


**Table `r iTab=iTab+1; iTab`.** Covariances from SS


```{r, covar}
cor =sscor(base$CoVar)
vars=c(paste("Bratio",2020:2022,sep="_"),paste("F",2020:2022,sep="_"))
cor =cor[vars,vars]
        
print(cor,digits=2)
```

**Table `r iTab=iTab+1; iTab`.** Covariances simulated by MVLN

```{r, covar-2, eval=FALSE}
cor=cor(cast(subset(as.data.frame(mvln[c("stock","harvest")],drop=T),year%in%2020:2022),iter~qname+year,value="data")[,-1])
dimnames(cor)=list(vars, vars) 

print(cor,digits=2)
```

**Table `r iTab=iTab+1; iTab`.** Estimates of $SSB/B_{MSY}$

```{r}
window(hatFLQ(sshat(bsh300$derived_quants))[c("stock","harvest")],start=2020,end=2023)
```

**Table `r iTab=iTab+1; iTab`.** Summary of MVLN terminal estimates of $SSB/B_{MSY}$

```{r, summary-ssb-mvln}
print(cast(as.data.frame(quantile(mvln[["stock"]][,ac(2020:2022)],probs=c(0.15,0.5))),iter~year,value="data"),digits=2)
```


**Table `r iTab=iTab+1; iTab`.** Summary of MVLN terminal estimates of $F/F_{MSY}$

```{r, summary-f-mvln}
print(cast(as.data.frame(quantile(mvln[["harvest"]][,ac(2020:2022)],probs=c(0.35,0.5))),iter~year,value="data"),digits=2)
```


```{r, k2sm}
k2sm=transform(ldply(mvls, function(x) if("try-error"%in%is(x)) return(NULL) else as.data.frame(x)),Run=runs[an(X1)])
k2sm=ddply(k2sm,  .(Run,year), with, kobe:::prob(stock,harvest))
k2sm=ddply(melt(k2sm,id=c("Run","year")), .(Run,variable,year), with, mean(value))

p=ggplot(subset(k2sm,variable=="green"&year<=2050))+
  geom_line(aes(year,V1,col=Run))
k2sm=cast(k2sm,variable~Run~year,value="V1","mean")[,,ac(1971:2070)] 
```

**Table `r iTab=iTab+1; iTab`.** Probability of being in green quadrant.

```{r, k2sm-f}
grn=k2sm["green",,]
grn[is.na(grn)]=0
grn[,-1]=paste(unlist(c(as.integer(grn[,-1]*100))),"%",sep="")
grn[,ac(2023:2030)]
```


**Table `r iTab=iTab+1; iTab`.** Probability of being over fished.

```{r, k2sm-ofd}
oFd=k2sm["overFished",,]

oFd[is.na(oFd)]=1
oFd[,-1]=paste(unlist(c(as.integer(oFd[,-1]*100))),"%",sep="")
oFd[,ac(2023:2030)]
```

**Table `r iTab=iTab+1; iTab`.** Probability of overfishing.

```{r, k2sm-ofg}
oFg=k2sm["overFishing",,]

oFg[is.na(oFg)]=1
oFg[,-1]=paste(unlist(c(as.integer(oFg[,-1]*100))),"%",sep="")
oFg[,ac(2023:2030)]
```

```{r}
library(kobe)
library(xtable) 
library(tables) 
library(Hmisc)

#\definecolor{gray50}{rgb}{ 1.0,1.0,1.0}
#\definecolor{gray60}{rgb}{ 0.9,0.9,0.9}
#\definecolor{gray70}{rgb}{ 0.8,0.8,0.8}
#\definecolor{gray80}{rgb}{ 0.7,0.7,0.7}
#\definecolor{gray90}{rgb}{ 0.6,0.6,0.6}
#\definecolor{gray100}{rgb}{0.5,0.5,0.5}

t1=latex(kobe:::shade(k2sm["green",,ac(2023:2032)],pct=""),
      file    ="../outputs/tex/k2smGreen.tex", 
      rowlabel="TAC",
      rowname =paste(dimnames(k2sm["green",,])[[1]],"t"),
      caption ="Kobe II Strategy Matrix, $P(F\\leq F_{MSY})$.")

t2=latex(kobe:::shade(k2sm["overFished",,ac(2023:2032)],pct=""),
      file    ="../outputs/tex/overfished.tex", 
      rowlabel="TAC",
      rowname =paste(dimnames(k2sm["overFished",,])[[1]],"t"),
      caption ="Kobe II Strategy Matrix, $P(F\\leq F_{MSY})$.")

t3=latex(kobe:::shade(k2sm["overFishing",,ac(2023:2032)],pct=""),
      file    ="../outputs/tex/overfishing.tex", 
      rowlabel="TAC",
      rowname =paste(dimnames(k2sm["overFishing",,])[[1]],"t"),
      caption ="Kobe II Strategy Matrix, $P(F\\leq F_{MSY})$.")
```

```{r, eval=FALSE}
kab=knitr::kable(k2sm["overFished",,ac(2023:2032)], caption="Table 1: ", booktabs=T, digits=2, row.names=FALSE)
kable_classic_2(kab, full_width=F, latex_options="hold_position")
```

\normalsize


```{r,echo=FALSE, eval=FALSE}
sim<-function(covar,hat,year,biomass="Bratio",f="F",ref="annF_MSY",fyr=-1,
              mc=5000,ss3diag=TRUE,puredelta=TRUE){

  y=rbind(hat[grep(paste(biomass,year,sep="_"),hat$Label),],
          hat[grep(paste(f,      year+fyr,  sep="_"),hat$Label),])
  fref = hat[grep(ref,hat$Label), ]
  
  bratio=covar[ grep(paste(biomass,year,sep="_"),covar$label.i),]
  x=bratio[grep(paste(f,year+fyr,sep="_"),bratio$label.j),]
  fs=covar[grep(ref,covar$label.i),]
  x1=fs[grep(paste(biomass,year,sep="_"),fs$label.j),]
  x2=fs[grep(paste(f,year+fyr,sep="_"),fs$label.j),]
  ss=rbind(x,x1,x2)
  
  
  ## Correlations on the log scale
  ## Estimates on the untransformed scale
  if (ss3diag){
  varF    = log(1 + (y$StdDev[1]/y$Value[1])^2)
  varB    = log(1 + (y$StdDev[2]/y$Value[2])^2)
  varFref = log(1 + (fref$StdDev[1]/fref$Value)^2)
  
  cov =  log(1 + mean(x$corr)*sqrt(varF*varB))
  cov1 = log(1 + mean(x1$corr)*sqrt(varB*varFref))
  cov2 = log(1 + mean(x2$corr)*sqrt(varF*varFref))
  
  mvnmu = log(c(y$Value[2], y$Value[1], fref$Value))
  
  mvncov = matrix(NA, ncol = 3, nrow = 3)
  diag(mvncov) = c(varB, varF, varFref)
  mvncov[1, 2] = mvncov[2, 1] = cov
  mvncov[2, 3] = mvncov[3, 2] = cov1
  mvncov[1, 3] = mvncov[3, 1] = cov2
  
  rtn=exp(mvtnorm::rmvnorm(mc, mean = mvnmu, sigma = mvncov, method = c("svd")))
  }else{
  varlogF   =(y$StdDev[1]/y$Value[1])^2
  varlogB   =(y$StdDev[2]/y$Value[2])^2
  varlogFref=(fref$StdDev[1]/fref$Value)^2
 
  if(!puredelta){
    varlogF = log(1 + varlogF)
    varlogB = log(1 + varlogB)
    varlogFref = log(1 + varlogFref)}
  
  cov  =mean(x$corr)*sqrt(varlogF*varlogB)
  cov1 =mean(x1$corr)*sqrt(varlogB*varlogFref)
  cov2 =mean(x2$corr)*sqrt(varlogF*varlogFref)
  
  if(!puredelta){
    cov = log(1 + cov)
    cov1 = log(1 + cov1)
    cov2 = log(1 + cov2)}
  
  mvncov = matrix(NA, ncol = 3, nrow = 3)
  diag(mvncov) = c(varlogB, varlogF, varlogFref)
  mvncov[1, 2] = mvncov[2, 1] = cov
  mvncov[2, 3] = mvncov[3, 2] = cov1
  mvncov[1, 3] = mvncov[3, 1] = cov2
  
  #MLE of log(Bratio), log(F), log(Fref) is obtained by direct transformation of MLE in non-log scale:
  mvnmu = log(c(y$Value[2], y$Value[1], fref$Value)) 
  
  rtn=exp(mvtnorm::rmvnorm(mc, mean = mvnmu, sigma = mvncov, method = c("svd")))}
  
  rtn=as.data.frame(rtn)
  names(rtn)=c("harvest","stock","f")
  
  list(sim=rtn,ss=ss)}

hat  =bsh$derived_quants
covar=bsh$CoVar

set.seed(87888)
mc1=sim(covar,hat,year=2021,mc=50000)
set.seed(87888)
mc2=sim(covar,hat,year=2021,mc=50000,ss3diag=FALSE,puredelta=!TRUE)
set.seed(87888)
mc3=sim(covar,hat,year=2021,mc=50000,ss3diag=FALSE,puredelta=TRUE)

mc1[[2]]
sims=rbind(cbind(run="SSMVLN",    mc1[[1]]),
           cbind(run="CF",        mc2[[1]]),
           cbind(run="Pure Delta",mc3[[1]]))


kobe:::kobePhaseMar2(sims,
                     ylim=2.5,xlim=2.5,
                     xlab=expression(B/B[MSY]),
                     ylab=expression(F/F[MSY]),
                     quadcol=c("red","green","yellow","orange"),size=0.1)

ddply(sims,.(run), with, quantile(stock))
ddply(sims,.(run), with, quantile(harvest))
ddply(sims,.(run), with, cor(log(harvest),log(stock)))

mc1[[2]]

nms =paste(c("Bratio","F"),rep(as.character(seq(2020,2022)),each=2),sep="_")
flag=bsh$CoVar[,7]%in%nms&bsh$CoVar[,8]%in%nms
cast(bsh$CoVar[flag,7:9],label.i~label.j)
```

```{r,eval=FALSE}
dat=read.csv("/home/laurie/Desktop/risk/caseStudies/bsh/ss_forecast/forecasts/150/forecast.ss",skip=63,sep="",header=F,col.names=c("year","v2","fleet","data"))
ggplot(subset(dat,year<2040))+geom_point(aes(year,data))+facet_grid(fleet~.,scale="free")
```


```{r, eval=FALSE}
library(calculus)
library(compositions)

sim1=exp(mvtnorm::rmvnorm(1000, mean=mvlmu[c("Bratio_2020","F_2020")], 
                               sigma=cvr[  c("Bratio_2020","F_2020"),
                                         c("Bratio_2020","F_2020")], 
                          method = c("svd")))
sim2=rlnorm.rplus(1000,mvlmu[c("Bratio_2020","F_2020")], cvr[c("Bratio_2020","F_2020"),c("Bratio_2020","F_2020")])

fn<-function(x,y)
 dlnorm.rplus(c(x,y),   mvlmu[c("Bratio_2020","F_2020")],
                        cvr[  c("Bratio_2020","F_2020"),
                              c("Bratio_2020","F_2020")])
integral(fn, bounds=list(x=c(0,10),y=c(-10,0)))$value


k2sm["green",1,"2020"]
cor=sscor(bsh$CoVar)[c("Bratio_2020","F_2020"),
                     c("Bratio_2020","F_2020")]
hat=subset(sshat(bsh$derived_quants),label%in%c("Bratio_2020","F_2020"))
cov=cor2cov(cor,hat[,"cv"])

fn<-function(x,y)
 dnorm.aplus(c(x,y),hat[,"hat"],cov)
integral(fn, bounds=list(x=c(1,0),y=c(Inf,1)))$value
  
```


