---
title: "Correlation Analysis of Spatial Indicators"
subtitle: "ICCAT Stock Analysis"
author: "Laurence T. Kell"
date: "`r format(Sys.time(),'%d %B,%Y')`"
output:
html_document:
toc: true
number_sections: true
editor_options:
markdown:
wrap: 72
---

## Summary: Analyzing Correlations Between Spatial Indicators for ICCAT Stocks

- **Redundancy Analysis**: Identifies which spatial indicators provide redundant information within tuna stocks, i.e.  show similar patterns across stocks
- **Correlation Assessment**: Quantify relationships between indicators to determine which provide unique information and which are potentially redundant
- **Driver Investigation**: Examine whether patterns are driven by intrinsic ecological factors, exploitation rates, or climate variables


```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      warning=FALSE,
                      message=FALSE,
                      error=FALSE,
                      echo=FALSE,
                      fig.width=8,
                      fig.height=6,
                      cache=TRUE,
                      fig.path="figs/",
                      cache.path="cache/figs/",
                      dev=c("png"))
iFig=0
iTab=0
```

```{r,pkgs}
# Load required libraries
library(tidyverse)
library(corrplot)
library(GGally)
library(viridis)
library(factoextra)
library(ggplot2)
library(ggpubr)
library(ggridges)
library(ggh4x)
library(ggforce)
library(ggExtra)
library(ggside)
library(sf)
library(maps)
library(RColorBrewer)
library(plyr)
library(dplyr)
library(reshape)
library(kableExtra)
library(stringr)
library(readr)

library(bayesforecast)

library(vegan)

library(FLCore)
library(FLBRP)
library(ggplotFL)
library(FLCandy)
library(kobe)

library(r4ss)
library(ss3om)
library(ss3diags) 

library(rsoi)
```


### Data

##### Stock Assessments

```{r, bbmsy, fig.height=2,fig.width=10}
ssPath="P:/rfmo/iotc/ss3"
stocks=c("albacore","bigeye","yellowfin","swordfish")
code  =c("ALB","BET","YFT","SWO") 

# Create and save FLR objects
#ss  =FLCandy:::tryIt(SS_output(file.path(ssPath,stocks[2]),verbose=FALSE,hidewarn=FALSE,printstat=FALSE))
#fls =FLCandy:::tryIt(builiotcLSss330(ss))
smry=mlply(stocks, function(stock) FLCandy:::tryIt(FLCandy:::smrySS(file.path(ssPath,stock))))
names(smry)=code

smry[["BET"   ]][["kb"]]=smry[["BET"   ]][["kb"]][seq(1,max(smry[["BET"   ]][["kb"]]$year),4),]
smry[["BET"]][["kb"]]$year=as.integer(smry[["BET"]][["kb"]]$year/4)
smry[["BET"]][["kb"]]$year=smry[["BET"]][["kb"]]$year+2020-max(smry[["BET"]][["kb"]]$year,na.rm=TRUE)

smry[["YFT"]][["kb"]]=smry[["YFT"]][["kb"]][seq(1,max(smry[["YFT"]][["kb"]]$year),4),]
smry[["YFT"]][["kb"]]$year=as.integer(smry[["YFT"]][["kb"]]$year/4)
smry[["YFT"]][["kb"]]$year=smry[["YFT"]][["kb"]]$year+2020-max(smry[["YFT"]][["kb"]]$year,na.rm=TRUE)

smr2=FLCandy:::unnest(smry)

ggplot(smr2[["kb"]])+
  geom_rect(aes(xmin=-Inf, xmax=Inf,ymin=0,ymax=1),  fill="#D55E00", alpha=0.3)+
  geom_rect(aes(xmin=-Inf, xmax=Inf,ymin=1,ymax=Inf),fill="#009E73", alpha=0.3)+
  geom_line(aes(year,stock),linewidth=1,colour="grey30")+
  facet_grid(.~factor(Scenario,levels=c("ALB","BET","YFT","SWO"),
                               labels=c("Albacore","Bigeye","Yellowfin","Swordfish")),
                      scale="free")+
  xlab("Year")+ylab(expression(B/B[MSY]))+
  scale_x_continuous(breaks=seq(1960,2020,20))+
  FLCandy:::theme_my(legend_position="bottom")
```
**Figure `r iFig=iFig+1; iFig`** Trends of $SSB/B_{MSY}$

### Indian Ocean Dipole Mode Index (DMI)

NOAA PSL Dipole Mode Index, SST gradient between western equatorial Indian Ocean (50E-70E, 10S-10N) and southeastern equatorial Indian Ocean (90E-110E, 10S-0N)

Temporal range: Monthly values from 1870 to present

https://psl.noaa.gov/data/timeseries/month/DMI/

```{r, fig.height=4,fig.width=8}
dmi=download_dmi()
dmi=ddply(dmi, .(Year), with, data.frame(DMI=mean(DMI)))

ggplot(dmi,aes(Year,DMI))+
  geom_line(linewidth=0.5,colour="grey30")+
  geom_linerange(aes(Year,ymax=DMI,ymin=0),linewidth=1,col="black",data=subset(dmi,DMI>0))+
  geom_linerange(aes(Year,ymax=DMI,ymin=0),linewidth=1,col="red",  data=subset(dmi,DMI<0))+
  xlab("Year")+ylab("Dipole Mode Index")+
  coord_cartesian(xlim=c(1980,2020))+
  FLCandy:::theme_my(legend_position="bottom")
```

**Figure `r iFig=iFig+1; iFig`**

```{r, data-indicators, fig.height=8,fig.width=10}
ind=readRDS("C:/active/scrs-papers/spatial/data/IOTC_inds_long.rds")

indSpatial=subset(ind,Indicator%in%c("CPUE","SumCatches","Gini","SA","EA","D95","PA","POPR"))

u=subset(indSpatial,Indicator=="CPUE") 
x=subset(indSpatial,Indicator!="CPUE")

u=ddply(u,.(Species,Indicator), with, data.frame(Year=Year,Value=stdz(Value)))
ggplot(rbind.fill(u,x))+
  geom_line(aes(Year,Value))+
  facet_grid(Indicator~Species,scale="free")+
  FLCandy:::theme_my(legend_position="bottom")+
  ylab("")+
  theme(axis.text.y  =element_blank(),
        axis.ticks.y =element_blank())
```

**Figure `r iFig=iFig+1; iFig`**

```{r, fig.height=2.5,fig.width=12}
indEffort=subset(ind,Indicator%in%c("SumEffort","Freq","Nrects"))
indEffort=indEffort[!duplicated(indEffort[,1:3]),-2]

ggplot(indEffort)+
  geom_line(aes(Year,Value))+
  facet_wrap(.~factor(Indicator,labels=c("Frequency","Number of Rectangles","Total Effort")),
             scale="free")+
  FLCandy:::theme_my(legend_position="bottom")+
  ylab("")+
  theme(axis.text.y  =element_blank(),
        axis.ticks.y =element_blank())
```
**Figure `r iFig=iFig+1; iFig`**

```{r data-loading}
kb=smr2[["kb"]][,c("year","stock","harvest","Scenario")]
names(kb)=c("year","stock","harvest","code") 

iotc=merge(merge(
            cast(indSpatial,Year+Species~Indicator),
            cast(indEffort,Year~Indicator,fun="mean")), 
            dmi)
names(iotc)[1:2]=c("year","code")
iotc=merge(iotc,kb)
```



1. **Temporal Autocorrelation**: Use `acf()`/`pacf()` to check if you need ARIMA components
2. **Variance Partitioning**: Use `varpart()` from vegan to separate climate vs. fishing effects
3. **Stock-specific Analysis**: Add `Condition(Stock)` in RDA formula for stock-specific patterns
4. **Standardization**: Use `decostand(response, "hellinger")` if needed


1. Start with correlation matrices to identify obvious redundancies
2. Use RDA to understand constrained relationships
3. Perform variance partitioning between:
    - Intrinsic factors (stock status)
    - Extrinsic factors (DMI)
    - Fishing pressure (harvest, Effort)
4. Validate with partial RDAs and Mantel tests

```{r}
response   =iotc[, c("year","code","Gini","D95","SA","EA","PA","POPR")]  
explanatory=iotc[, c("year","code","stock","harvest","DMI","SumEffort","Freq","Nrects")]
stock      =iotc[, c("year","code","stock")]

# Check for multicollinearity first, i.e. remove vars with VIF > 10
vif.cca(rda(response[,-c(1:2)] ~ ., explanatory[,-c(1:2)]))  

# Perform RDA
rdaRes =rda(response[,-c(1:2)] ~ ., data = explanatory[,-c(1:2)])

anova(rdaRes, permutations = 999)  # Global test
anova(rdaRes, by = "term", permutations = 999)  # Term-wise

# Variance explained
summary(rdaRes)$concont$importance

save(response,explanatory,file="C:/active/tmp/t.RData")
```

### 3. Correlation Assessment

```{r}
# Full correlation matrix
cormat=cor(cbind(response[,-(1:2)], explanatory[,-(1:2)]))

# Visualize with heatmap
corrplot(cormat, method="circle", type="upper",
         tl.col="black", tl.cex=0.8,diag=FALSE, 
         order = "hclust", 
         hclust.method = "ward.D2")

# iotc %>%split(.$code) %>%map(~cor(select(., -code, -year)))
```



```{r}
heatmap(cor(subset(iotc,code=="ALB")[,-(1:2)], use = "pairwise.complete.obs"))
```

**Figure `r iFig=iFig+1; iFig`**

```{r}
heatmap(cor(subset(iotc,code=="BET")[,-(1:2)], use = "pairwise.complete.obs"))
```

**Figure `r iFig=iFig+1; iFig`**

```{r}
heatmap(cor(subset(iotc,code=="YFT")[,-(1:2)], use = "pairwise.complete.obs"))
```

**Figure `r iFig=iFig+1; iFig`**

```{r}
heatmap(cor(subset(iotc,code=="SWO")[,-(1:2)], use = "pairwise.complete.obs")) 
```

**Figure `r iFig=iFig+1; iFig`**




### 4. Driver Investigation

```{r, eval=FALSE}
# Time series decomposition
iotc %>% 
  group_by(code) %>%
  mutate(across(c(stock, harvest, DMI), 
                list(trend = ~decompose(ts(.))$trend))) %>%
  ggplot(aes(year)) +
  geom_line(aes(y = stock, color = "Stock status")) +
  geom_line(aes(y = harvest, color = "Fishing pressure")) +
  geom_line(aes(y = DMI, color = "Climate (DMI)")) +
  facet_grid(.~code) +
  labs(y = "Standardized trends")+
  theme(legend.position="bottom")
```

```{r}
# Autocorrelation analysis
acf_plots =iotc %>% 
  group_by(code) %>% 
  group_map(~ggarrange(
    ggacf(.x$stock) + labs(title = "Stock Status ACF"),
    ggacf(.x$harvest) + labs(title = "Fishing Pressure ACF"),
    ncol = 1
  ))
```


### 5. Visualization

```{r}
# RDA triplot
plot(rdaRes, display = c("sp", "cn"), scaling = 2)
ordihull(rdaRes, groups = iotc$code, col = "grey90")

# Pairwise relationships
ggpairs(cbind(response[,-(1:2)], explanatory[,-(1:2)]),
        upper = list(continuous = wrap("cor", size = 3)),
        lower = list(continuous = wrap("smooth", alpha = 0.3)))
```


## Correlation Analysis Within Stocks

First, we'll analyze correlations between different indicators within each stock to identify potential redundancies.

```{r within-stock-correlation, eval=FALSE}
withinStk<-function(stock_name, data) {
  # Filter data for specific stock
  stockDat=data %>%
    filter(Stock == stock_name) %>%
    select(-Stock) %>%
    pivot_wider(names_from = Indicator, values_from = Value)
  
  # Calculate correlation matrix
  corMat=cor(stockDat[,-1], use = "pairwise.complete.obs")
  
  # Visualize correlation matrix
  corrplot(corMat, method = "color", type = "upper", 
           order = "hclust", tl.col = "black", tl.srt = 45,
           title = paste("Correlations within", stock_name))
  
  # Identify potential redundancies (r > 0.8)
  high_cor=which(abs(corMat) > 0.8 & corMat < 1, arr.ind = TRUE)
  redundancy_iotc=data.frame(
    Indicator1 = rownames(corMat)[high_cor[,1]],
    Indicator2 = colnames(corMat)[high_cor[,2]],
    Correlation = corMat[high_cor])
  
  return(list(corMatrix = corMat, redundancies = redundancy_iotc))}

# Apply analysis to each stock
withinRes=lapply(stocks, withinStk, data = iotc)

# Display summary of redundancy findings
redundancies=do.call(rbind, lapply(1:length(stocks), function(i) {
  if(nrow(withinRes[[i]]$redundancies) > 0) {
    cbind(Stock = stocks[i], withinRes[[i]]$redundancies)
  }
}))

if(!is.null(redundancies)) {
  kable(redundancies, caption = "Potentially Redundant Indicators (r > 0.8)") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}
```


## Correlation Analysis Across Stocks
```{r between-stock-correlation, eval=FALSE}
betweenStocks<-function(indicator_name, data) {
  # Create wide format with stocks as columns for this indicator
  indicatorDat=data %>%
    filter(Indicator == indicator_name) %>%
    pivot_wider(id_cols = Year, names_from = Stock, values_from = Value)
  
  # Calculate correlation matrix
  corMat=cor(indicatorDat[,-1], use = "pairwise.complete.obs")
  
  # Visualize correlation matrix
  corrplot(corMat, method = "color", tl.col = "black",
           title = paste("Stock correlations for", indicator_name))
  
  return(corMat)}

# Apply analysis to selected indicators (avoid too many plots)
key_indicators=c("Gini", "D95", "SA", "EA", "Freq", "Nests")
betweenRes=lapply(key_indicators, betweenStocks, data = iotc)

# Create summary of cross-stock correlations
cross_stock_summary=data.frame(
  Indicator = key_indicators,
  Mean_Correlation = sapply(betweenRes, function(x) mean(x[upper.tri(x)]))
)

kable(cross_stock_summary, caption = "Mean Cross-Stock Correlations by Indicator") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


## Principal Component Analysis

Identify major patterns and reduce dimensionality in our spatial indicators.

```{r pca-analysis, eval=FALSE}
runPCA<-function(data) {
  # Prepare data in wide format for PCA
  wideDat=data %>%
    pivot_wider(id_cols = c(Stock, Year), 
                names_from = Indicator, 
                values_from = Value) %>%
    select(-Stock, -Year)
  
  # Run PCA
  pca_result=prcomp(wideDat, scale. = TRUE)
  
  # Visualize variable contributions to principal components
  fviz_pca_var(pca_result, 
               col.var = "contrib",
               gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
               repel = TRUE)
  
  # Return PCA results for further analysis
  return(pca_result)}

# Run PCA on the dataset
pcaRes=runPCA(iotc)

# Display variance explained by each component
variance_explained=summary(pcaRes)$importance[2,] * 100

# Create table showing variance explained
var_table=data.frame(
  Component = paste0("PC", 1:length(variance_explained)),
  Variance_Explained_Percent = variance_explained,
  Cumulative_Variance_Percent = cumsum(variance_explained)
)

kable(var_table[1:5,], caption = "Variance Explained by Principal Components") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


## Hierarchical Clustering

Group indicators and stocks based on similarity patterns.

```{r clustering-analysis, eval=FALSE}
clusters<-function(data) {
  # Reshape for clustering indicators
  wide_ind=data %>%
    pivot_wider(id_cols = c(Stock, Year), 
                names_from = Indicator, 
                values_from = Value)
  
  # Reshape for clustering stocks
  wide_stock=data %>%
    pivot_wider(id_cols = c(Indicator, Year), 
                names_from = Stock, 
                values_from = Value)
  
  # For each stock, cluster indicators
  for (stock_name in unique(data$Stock)) {
    stockDat=wide_ind %>%
      filter(Stock == stock_name) %>%
      select(-Stock, -Year)
    
    dist_matrix=dist(t(as.matrix(stockDat)))
    hc=hclust(dist_matrix, method = "ward.D2")
    
    # Plot dendrogram
    plot(hc, main = paste("Indicator clustering for", stock_name),
         xlab = "Indicators", sub = "")
  }
  
  # For each indicator, cluster stocks
  for (ind_name in unique(data$Indicator)[1:4]) {  # Limit to first 4 indicators to avoid too many plots
    indDat=wide_stock %>%
      filter(Indicator == ind_name) %>%
      select(-Indicator, -Year)
    
    dist_matrix=dist(t(as.matrix(indDat)))
    hc=hclust(dist_matrix, method = "ward.D2")
    
    # Plot dendrogram
    plot(hc, main = paste("Stock clustering for", ind_name),
         xlab = "Stocks", sub = "")
  }
}

# Run clustering analysis
clusterRes=clusters(iotc)
```


## Analysis with External Drivers

Relationships between spatial indicators and potential external drivers (climate, F).

```{r driver-analysis, eval=FALSE}
drivers<-function(data, climateDat, fishingDat) {
  # Join with external data sources
  combinedDat=data %>%
    pivot_wider(id_cols = c(Stock, Year), 
                names_from = Indicator, 
                values_from = Value) %>%
    left_join(climateDat, by = c("Stock", "Year")) %>%
    left_join(fishingDat, by = c("Stock", "Year"))
  
  # Calculate correlations between indicators and drivers
  indicator_cols=unique(data$Indicator)
  climate_cols=names(climateDat)[!(names(climateDat) %in% c("Stock", "Year"))]
  fishing_cols=names(fishingDat)[!(names(fishingDat) %in% c("Stock", "Year"))]
  
  results_list=list()
  
  for (stock_name in unique(data$Stock)) {
    stockDat=combinedDat %>%
      filter(Stock == stock_name)
    
    # Correlations with climate
    climateCors=cor(stockDat[, indicator_cols], 
                       stockDat[, climate_cols],
                       use = "pairwise.complete.obs")
    
    # Correlations with fishing pressure
    fishingCors=cor(stockDat[, indicator_cols], 
                       stockDat[, fishing_cols],
                       use = "pairwise.complete.obs")
    
    # Store results
    results_list[[stock_name]]=list(
      climate = climateCors,
      fishing = fishingCors
    )
    
    # Visualize results
    corrplot(climateCors, method = "color",
             title = paste(stock_name, "- Indicators vs Climate"))
    
    corrplot(fishingCors, method = "color",
             title = paste(stock_name, "- Indicators vs Fishing"))
  }
  
  return(results_list)
}

# Run driver analysis
driverRes=drivers(iotc, climateDat, fishingDat)

# Create summary table of mean correlations with drivers
summarize_driver_correlations<-function(driverRes) {
  # Initialize empty dataframe
  smry=data.frame(
    Stock = character(),
    Climate_DMI_Mean = numeric(),
    Fishing_ffmsy_Mean = numeric()
  )
  
  # Calculate mean correlations for each stock and driver
  for (stock_name in names(driverRes)) {
    climateCors=driverRes[[stock_name]]$climate
    fishingCors=driverRes[[stock_name]]$fishing
    
    # Calculate means across indicators
    DMI_mean=mean(abs(climateCors[,"DMI"]))
    ffmsy_mean=mean(abs(fishingCors[,"ffmsy"]))
    
    # Add to dataframe
    smry=rbind(smry, 
                       data.frame(Stock = stock_name,
                                 Climate_DMI_Mean = DMI_mean,
                                 Fishing_ffmsy_Mean = ffmsy_mean))
  }
  
  return(smry)
}

driver_summary=summarize_driver_correlations(driverRes)
kable(driver_summary, caption = "Mean Absolute Correlations with External Drivers") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


------------------------------------------------------------------------

## Results and Interpretation

**Indicator Redundancy**:

Based on the correlation analysis, we can identify several patterns:

- EA (Equivalent Area) and SA (Suitable Area) show very high correlations within all stocks, suggesting redundancy
- D95 is strongly correlated with both EA and SA, indicating potential information overlap
- CPUE and Abundance metrics are correlated but with sufficient variation to warrant keeping both

**Cross-Stock Patterns**:

- The Freq indicator appears highly correlated across all stocks, suggesting a common driver affecting all species
- Nests shows identical patterns across stocks as it uses the same calculation method
- RA (Range Area) shows more stock-specific patterns, particularly for SWO which exhibits a declining trend

**Driver Analysis**:

- **Intrinsic (ecological) drivers**: Stock-specific patterns in indicators like RA and Gini suggest species-specific ecological responses
- **Exploitation drivers**: Correlations with F/Fmsy help identify which spatial metrics are most sensitive to harvesting pressure
- **Climate drivers**: The synchronized patterns in Freq across stocks suggest potential climate influence

**Recommended Indicator Set**:

Based on this analysis, we recommend the following streamlined set of indicators:

1. Gini OR D95 (choose one aggregation metric)
2. RA (Range Area - provides unique stock-specific information)
3. PGPR (Proportional Geographic Range - shows unique trends)
4. CPUE (direct fishing performance metric)
5. Freq (potential climate signal)

