---
title: "Spatial Indicators"
subtitle: "ICCAT"
author: 
  L.T. Kell
date: "`r format(Sys.time(),'%d %B,%Y')`"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Estimating $F_{MNY}$ for Atlantic Mackerel}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

## Summary: Using Spatial Indicators with ICCAT Data for Ecosystem Assessment and BBNJ Alignment


- ICCAT’s Ecosystem Approach: ICCAT is moving toward ecosystem-based fisheries management, aligning with international frameworks like the BBNJ Agreement and SEDA-SAP, which emphasize the use of the best available science, spatial management, and resilience-building.
- Spatial Indicators: Empirical spatial indicators (e.g., Gini, D95, occupancy) are increasingly recognized as valuable proxies for stock status, especially for data-limited stocks, and can describe spatial distribution, aggregation, and occupancy-key.


```{r,eval=FALSE,echo=FALSE}
setwd("C:/active/scrs-papers/spatial/Rmd")
```


```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE)

library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE,
               fig.width =8, 
               fig.height=6,
               cache     =TRUE, 
               fig.path  ="figs/",
               cache.path="cache/figs/",
               dev        =c("png"))

iFig=0
iTab=0
```

```{r,pkgs}
require(ggplot2)
require(ggpubr)
require(ggplotFL)
library(ggridges)
library(ggplot2)
library(ggh4x)
library(ggforce)
library(ggExtra)
library(ggside)

library(sf)
library(maps)
library(viridis)
library(RColorBrewer)

library(plyr)
library(dplyr)
library(reshape)
library(kableExtra)

library(stringr)

library(readr)

library(kableExtra)
```


### Data Preparation

-   Datasets:
    -   EFFDIS: Longline fishing effort (hooks) by 5°×5° cell, year, month, flag, gear.
    -   CATDIS: Catch (tons) by 5°×5° cell, year, trimester, flag, gear, species.
-   Merging:
    -   Align by year (or convert month to trimester), grid cell (Lat5/Lon5), flag, gear.
    -   Calculate CPUE: $\text{CPUE} = \frac{\text{Catch (t)} \times 1000}{\text{Effort (hooks)}}$.


```{r}
effdis=read.csv("P:/rfmo/iccat/data/effdis_ll.1950.2015.csv")
catdis=read.csv("P:/rfmo/iccat/data/cdis5023-all9sp.csv")

# 1. First, convert month to trimester in the EFFDIS dataset
effdis$Trimester=ceiling(effdis$month/3)

# 2. Create spatial grid identifiers that match between datasets
catdis$Lat5=round(catdis$yLat5ctoid/5)*5
catdis$Lon5=round(catdis$xLon5ctoid/5)*5
effdis$Lat5=round(effdis$latitude/5)*5
effdis$Lon5=round(effdis$longitude/5)*5

# 3. Ensure consistent naming conventions
names(effdis)[names(effdis)=="flagname"]="FlagName"
names(effdis)[names(effdis)=="year"]="YearC"
names(effdis)[names(effdis)=="geargrp"]="GearGrp"

# 4. Merge datasets on common fields
cpue=merge(catdis, effdis,by=c("YearC", "Trimester", "Lat5", "Lon5", "FlagName", "GearGrp"),
           all=FALSE)
cpue$stock=paste(cpue$SpeciesCode,cpue$Stock)

# 5. Calculate CPUE (tons per 1000 hooks)
cpue$cpue=(cpue$Catch_t * 1000)/cpue$hooksEst

iccat<-read.csv("C:/active/scrs-papers/spatial/data/Copy of ICCAT_B_Bmsy_and_F_Fmsy_ratios_forR_v4.csv")

key=c("AT","ATW","ATS","ATN","ATE")
names(key)=c("Atlantic","West Atlantic","South Atlantic","North Atlantic","East Atlantic")

iccat=transform(iccat,Stock=key[Stock])
prds =ddply(iccat, .(Year), with, ecdf(Value)(c(0.8,1.2)))

ggplot(prds)+geom_ribbon(aes(Year,ymin=0, ymax=V1),fill="red")+
             geom_ribbon(aes(Year,ymin=V1,ymax=V2),fill="green")+
             geom_ribbon(aes(Year,ymin=V2,ymax=1 ),fill="blue")+
  ylab("State relative to FAO criteria")+
  scale_x_continuous(limits=c(1950,NA))
```

```{r}
# Convert to presence/absence format for rectangle-level occupancy
presence_data=cpue %>%
  mutate(presence = ifelse(cpue > 0, 1, 0)) %>%
  group_by(YearC, stock) %>%
  summarize(
    totalRectangles = n(),
    occupiedRectangles = sum(presence),
    POPR = occupiedRectangles/totalRectangles  # Similar to paRect() function
  )
```

```{r}
# Calculate Gini index for each species/year
giniResults=cpue %>%
  group_by(YearC, stock) %>%
  group_modify(~{
    # Sort CPUE values in ascending order
    sorted_cpue=sort(.x$cpue)
    n=length(sorted_cpue)
    
    # Calculate cumulative proportions
    cum_prop=cumsum(sorted_cpue)/sum(sorted_cpue)
    
    # Calculate Gini index (similar to Gini() function)
    gini_value=1 - (2 * sum(cum_prop) / n)
    
    data.frame(gini = gini_value)
  })

```

```{r}

# Calculate D95 index (area containing 95% of CPUE)
d95Results=cpue %>%
  group_by(YearC, stock) %>%
  group_modify(~{
    # Sort CPUE values in descending order
    sorted_data=.x[order(-.x$cpue),]
    
    # Calculate cumulative proportion of CPUE
    sorted_data$cum_cpue=cumsum(sorted_data$cpue)/sum(sorted_data$cpue)
    
    # Find minimum area containing 95% of CPUE
    d95_value=min(which(sorted_data$cum_cpue >= 0.95))/nrow(sorted_data)
    
    data.frame(d95 = d95_value)
  })
```

```{r}
# Calculate Center of Gravity (CoG) and Inertia
cog_inertia=cpue %>%
  group_by(YearC, stock) %>%
  summarize(
    # Center of Gravity (weighted mean position)
    cog_lat = weighted.mean(Lat5, w = cpue, na.rm = TRUE),
    cog_lon = weighted.mean(Lon5, w = cpue, na.rm = TRUE),
    
    # Inertia (dispersion around CoG)
    inertia_lat = weighted.mean((Lat5 - weighted.mean(Lat5, w = cpue, na.rm = TRUE))^2, 
                                w = cpue, na.rm = TRUE),
    inertia_lon = weighted.mean((Lon5 - weighted.mean(Lon5, w = cpue, na.rm = TRUE))^2, 
                                w = cpue, na.rm = TRUE),
    inertia = inertia_lat + inertia_lon
  )
```

```{r}
# 7. SUMMARIZE BY SPECIES, YEAR, AND LOCATION
cpueSmry<-cpue %>%
  group_by(YearC, FlagName, stock, Lat5, Lon5) %>%
  summarize(
    total_catch = sum(Catch_t, na.rm = TRUE),
    total_effort = sum(hooksEst, na.rm = TRUE),
    cpue = (total_catch * 1000) / total_effort,
    .groups = "drop"
  )

# 8. CALCULATE SPATIAL INDICATORS

# 8.1 GINI INDEX CALCULATION
calculate_gini<-function(values) {
  # Sort values in ascending order
  sorted_values=sort(values)
  n=length(sorted_values)
  
  # Calculate cumulative proportions
  cum_prop=cumsum(sorted_values)/sum(sorted_values)
  
  # Calculate Gini index
  gini_value=1 - (2 * sum(cum_prop) / n)
  
  return(gini_value)
}

# 8.2 D95 INDEX CALCULATION 
calculate_d95<-function(values) {
  # Sort values in descending order
  sorted_data=sort(values, decreasing = TRUE)
  
  # Calculate cumulative proportion
  cum_sum=cumsum(sorted_data)
  cum_prop=cum_sum / sum(sorted_data)
  
  # Find minimum proportion containing 95% of values
  d95_value=min(which(cum_prop >= 0.95)) / length(sorted_data)
  
  return(d95_value)
}

# 8.3 CENTER OF GRAVITY CALCULATION
calculate_cog<-function(data) {
  # Calculate weighted means
  cog_lat=weighted.mean(data$Lat5, w = data$cpue, na.rm = TRUE)
  cog_lon=weighted.mean(data$Lon5, w = data$cpue, na.rm = TRUE)
  
  return(data.frame(cog_lat = cog_lat, cog_lon = cog_lon))
}

# 9. APPLY SPATIAL INDICATORS BY SPECIES AND YEAR
spatial_indicators=cpueSmry %>%
  group_by(YearC, stock) %>%
  filter(n() >= 5) %>% # Ensure enough cells for reliable indicators
  summarize(
    gini_index = calculate_gini(cpue),
    d95 = calculate_d95(cpue),
    cog = list(calculate_cog(cur_data())),
    n_cells = n(),
    mean_cpue = mean(cpue, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  tidyr::unnest_wider(cog)
```


------------------------------------------------------------------------

### Results

-   Aggregation Indicators (robust to coarse 5°×5° data):
    -   Gini Index: Measures how evenly catch/CPUE is distributed across space. Lower values = more aggregation; higher = more uniformity.
    -   D95: Proportion of area (cells) containing 95% of the catch/CPUE. Lower = high aggregation; higher = more spread.
-   Other Indicators (less robust with coarse data):
    -   Occupancy (POPR/POPH): Proportion of cells/hauls with presence; less informative with few cells.
    -   Center of Gravity: Mean location, can still be used for tracking shifts.

-   Gini & D95:
    -   High Gini/High D95: Evenly distributed stock, likely healthy or rebuilding.
    -   Low Gini/Low D95: Highly aggregated, potentially depleted or at risk.
    -   Trends: Declining Gini/D95 may signal contraction before biomass declines.
-   Limitations: With 5°×5° data, occupancy and fine-scale dispersion indicators are less reliable; focus on aggregation and location metrics.

```{r, fig.height=16,fig.width=12, eval=FALSE}
# 10.1 GINI AND D95 TIME SERIES PLOT
ggplot(spatial_indicators, aes(x = YearC)) +
  geom_point(aes(y = gini_index, color = "Gini Index"), size = 2) +
  geom_line(aes(y = gini_index, color = "Gini Index"), size = 1) +
  geom_point(aes(y = d95, color = "D95"), size = 2) +
  geom_line(aes(y = d95, color = "D95"), size = 1) +
  facet_wrap(~stock,ncol=2) +
  scale_color_manual(values = c("Gini Index" = "#1B9E77", "D95" = "#D95F02")) +
  labs(
    title = "Gini Index and D95 Trends by Species",
    #subtitle = "Higher values indicate more uniform distribution",
    x = "Year",
    y = "Indicator Value",
    color = "Indicator"
  ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "gray90", color = NA),
    strip.text = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11, color = "gray30")
  )
```

```{r, fig.height=8,fig.width=10}
# 10.2 GINI VS D95 SCATTERPLOT WITH SPECIES COMPARISON
ggplot(spatial_indicators, aes(x = gini_index, y = d95, color = stock)) +
  geom_point(aes(size = mean_cpue), alpha = 0.7) +
  scale_color_viridis_d() +
  labs(
    title = "Relationship Between Gini Index and D95 by Species",
    subtitle = "Point size indicates mean CPUE",
    x = "Gini Index (higher = more aggregated)",
    y = "D95 (higher = more spread)",
    color = "Species",
    size = "Mean CPUE"
  )+
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11, color = "gray30")
  ) +
  # Reference lines at median values
  geom_hline(yintercept = median(spatial_indicators$d95, na.rm = TRUE), 
             linetype = "dashed", color = "gray50", alpha = 0.7) +
  geom_vline(xintercept = median(spatial_indicators$gini_index, na.rm = TRUE), 
             linetype = "dashed", color = "gray50", alpha = 0.7)

# 10.3 CENTER OF GRAVITY MAP VISUALIZATION
# Create world map background
world_map=map_data("world")
```

```{r, fig.height=12,fig.width=12, eval=FALSE}
# Plot centers of gravity by species over time
ggplot() +
  geom_polygon(data=world_map, aes(x=long, y=lat, group=group), 
               fill="gray90", color="gray70", size=0.2) +
  geom_point(data=spatial_indicators, 
             aes(x=cog_lon, y=cog_lat, color=YearC), 
             size=2, alpha=0.8) +
  geom_path(data=spatial_indicators, 
             aes(x=cog_lon, y=cog_lat, color=YearC), 
             size=.25) +
  facet_wrap(~stock) +
  scale_color_viridis_c() +
  coord_quickmap(xlim=c(-100, 30), ylim=c(-20, 70)) +
  labs(
    title="Center of Gravity by Species and Year",
    subtitle="Color indicates year",
    x="Longitude",
    y="Latitude",
    color="Year"
  ) +
  theme_minimal() +
  theme(
    panel.grid=element_blank(),
    strip.background=element_rect(fill="gray90", color=NA),
    strip.text=element_text(face="bold"),
    axis.title=element_text(face="bold"),
    plot.title=element_text(face="bold", size=14),
    plot.subtitle=element_text(size=11, color="gray30") )

```
