---
title: "WKBSEABASS North"
subtitle: "Sensitivity Analysis for M and Steepness"
author: 
  L.T. Kell
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Estimating $F_{MNY}$ for Atlantic Mackerel}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---


# Implementing Spence's Ensemble Framework for Stock Assessment Models

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                     fig.width = 10, fig.height = 6)

# Load required packages
library(dplyr)
library(ggplot2)
library(tidyverse)
library(rstan)
library(reshape2)
library(mvtnorm)
library(grid)
library(gridExtra)

# Set Stan options for faster sampling
options(mc.cores = parallel::detectCores() - 2)
rstan_options(auto_write = TRUE)
```


## Introduction

In fisheries stock assessment, multiple models with different assumptions often yield diverse results. Rather than selecting a single "best" model, ensemble approaches combine information from multiple models to provide more robust estimates with better-characterized uncertainty. This document implements Michael A. Spence's discrepancy-based ensemble framework, which explicitly models the relationship between assessment outputs and the true state through discrepancy terms.

## Theoretical Background

Spence's approach recognizes that all models have discrepancies from the true state of the ecosystem. According to Spence's framework, the logarithm of each assessment's estimated spawning stock biomass (SSB) relates to the true SSB through the equation:

$
\ln(\hat{x}_k(t)) = \ln(y(t)) + \zeta_k(t) + \epsilon_k(t)
$

Where:

- $\hat{x}_k(t)$ is the estimated SSB from model $k$ at time $t$
- $y(t)$ is the true SSB at time $t$ (unknown)
- $\zeta_k(t)$ is the discrepancy of model $k$
- $\epsilon_k(t)$ represents parameter uncertainty, typically modeled as $N(0, \sigma^2_k(t))$

The discrepancy term can be further decomposed:

$
\zeta_k(t) = \gamma_k + \delta_k(t)
$

Where:

- $\gamma_k$ is the long-term discrepancy (fixed in time)
- $\delta_k(t)$ is the short-term discrepancy (changing over time)

Since assessment models often share similar structures and are fitted to the same data, their short-term discrepancies are typically correlated:

$
\delta_k(t) = \eta(t) + z_k(t)
$

Where:

- $\eta(t)$ is the shared short-term discrepancy across models
- $z_k(t)$ is the individual short-term discrepancy specific to model $k$


## Implementation

This document implements the ensemble approach for a grid of stock assessment models with different values for natural mortality (M) and steepness (h).

### Define Path and Scenarios

```{r scenarios}
# Define the working directory - modify as needed
path <- getwd()

# Define the grid of scenarios
scen <- expand.grid(M = seq(0.4, 1.2, 0.1),
                   h = c(0.8, 0.9, 0.999))
```


### Extract Model Outputs

```{r extract_data}
# Function to extract time series from stock assessment models
getSS <- function(M, h, path = getwd()) {
  # Load the saved SS output
  load(file.path(path, M, h, "ss.RData"))
  
  # Extract SSB time series with uncertainty
  ts <- ss$timeseries %>%
    dplyr::select(Yr, SpawnBio) %>%
    dplyr::rename(year = Yr, ssb = SpawnBio)
  
  # Extract uncertainty if available
  if ("derived_quants" %in% names(ss)) {
    sd <- ss$derived_quants %>%
      filter(grepl("^SSB_", Label)) %>%
      mutate(year = as.numeric(gsub("SSB_", "", Label))) %>%
      dplyr::select(year, StdDev) %>%
      dplyr::rename(ssb_sd = StdDev)
    
    ts <- left_join(ts, sd, by = "year")
  } else {
    # If standard deviation not available, use 10% of SSB as a rough estimate
    ts$ssb_sd <- ts$ssb * 0.1
  }
  
  # Add model identifiers
  ts$M <- M
  ts$h <- h
  
  return(ts)
}

# Apply function to all models in the grid
# Note: This is commented out as it requires the actual model outputs to be in place
# In practice, you would uncomment and run this
# outputs <- list()
# for(i in 1:nrow(scen)) {
#   M <- scen[i, "M"]
#   h <- scen[i, "h"]
#   outputs[[paste0("M", M, "_h", h)]] <- getSS(M, h, path)
# }
# 
# outputs <- bind_rows(outputs, .id = "model_id")
# outputs <- outputs %>%
#   mutate(logSSB = log(ssb),
#          logSD = ssb_sd / ssb)  # Delta method for SD on log scale
# 
# # Filter to a common time period across all models
# yrs <- outputs %>%
#   group_by(year) %>%
#   dplyr::summarize(count = dplyr::n()) %>%
#   filter(count == nrow(scen)) %>%
#   pull(year)
# 
# outputs <- outputs %>% 
#   filter(year %in% yrs) %>% 
#   filter(!is.na(ssb_sd))
```


### Define the Stan Model for Discrepancy-Based Ensemble

```{r stan_model}
stanCode <- "
data {
  int<lower=1> nObs;         // Total number of observations
  int<lower=1> nYrs;         // Number of years
  int<lower=1> nModels;      // Number of models
  
  int<lower=1> yearIdx[nObs]; // Year index for each observation
  int<lower=1> modelIdx[nObs]; // Model index for each observation
  
  vector[nObs] logSSB;        // Log SSB observations
  vector<lower=0>[nObs] logSD; // Observation standard deviations
}

parameters {
  // True state
  vector[nYrs] log_true_ssb;
  
  // Long-term discrepancy (gamma)
  vector[nModels] gamma;
  
  // Parameters for shared short-term discrepancy (eta)
  vector[nYrs] eta;
  real<lower=-1,upper=1> rho_eta;
  real<lower=0> sigma_eta;
  
  // Parameters for individual short-term discrepancy (z)
  matrix[nYrs, nModels] z;
  vector<lower=-1,upper=1>[nModels] rho_z;
  vector<lower=0>[nModels] sigma_z;
  
  // Process error for true state
  real<lower=-1,upper=1> rho_true;
  real<lower=0> sigma_true;
}

model {
  // Priors for true state process
  rho_true ~ uniform(-0.9, 0.9);
  sigma_true ~ cauchy(0, 1);
  
  // Prior for initial state
  log_true_ssb[^1] ~ normal(10, 2);
  
  // True state process (AR1)
  for (t in 2:nYrs) {
    log_true_ssb[t] ~ normal(rho_true * log_true_ssb[t-1], sigma_true);
  }
  
  // Priors for long-term discrepancy
  gamma ~ normal(0, 1);
  
  // Priors for shared short-term discrepancy
  rho_eta ~ uniform(-0.9, 0.9);
  sigma_eta ~ cauchy(0, 1);
  eta[^1] ~ normal(0, sigma_eta / sqrt(1 - square(rho_eta)));
  
  // Shared discrepancy process (AR1)
  for (t in 2:nYrs) {
    eta[t] ~ normal(rho_eta * eta[t-1], sigma_eta);
  }
  
  // Priors for individual short-term discrepancy
  rho_z ~ uniform(-0.9, 0.9);
  sigma_z ~ cauchy(0, 1);
  
  // Individual discrepancy processes (AR1)
  for (k in 1:nModels) {
    z[1, k] ~ normal(0, sigma_z[k] / sqrt(1 - square(rho_z[k])));
    for (t in 2:nYrs) {
      z[t, k] ~ normal(rho_z[k] * z[t-1, k], sigma_z[k]);
    }
  }
  
  // Likelihood
  for (i in 1:nObs) {
    int t = yearIdx[i];
    int k = modelIdx[i];
    
    logSSB[i] ~ normal(log_true_ssb[t] + gamma[k] + eta[t] + z[t, k], logSD[i]);
  }
}

generated quantities {
  vector[nObs] logSSBHat;
  vector[nObs] ssbHat;
  vector[nYrs] true_ssb;
  
  for (t in 1:nYrs) {
    true_ssb[t] = exp(log_true_ssb[t]);
  }
  
  for (i in 1:nObs) {
    int t = yearIdx[i];
    int k = modelIdx[i];
    logSSBHat[i] = log_true_ssb[t] + gamma[k] + eta[t] + z[t, k];
    ssbHat[i] = exp(logSSBHat[i]);
  }
}
"

# Save the Stan model
writeLines(stanCode, "discrepancy_model.stan")
```


### Prepare Data and Fit the Model

```{r fit_model, eval=FALSE}
# Prepare data for Stan
stan_data <- list(
  nObs = nrow(outputs),
  nYrs = length(yrs),
  nModels = nrow(scen),
  
  yearIdx = match(outputs$year, yrs),
  modelIdx = as.numeric(factor(outputs$model_id)),
  
  logSSB = outputs$logSSB,
  logSD = outputs$logSD
)

# Fit the model
ensembleFit <- stan(
  file = "discrepancy_model.stan",
  data = stan_data,
  iter = 2000,
  warmup = 1000,
  chains = 4,
  refresh = 200,
  control = list(adapt_delta = 0.9, max_treedepth = 12)
)

# Save the fitted model
save(ensembleFit, file = "ensemble_fit.RData")
```


### Extract Results and Create Summary Statistics

```{r extract_results, eval=FALSE}
# Extract posterior samples
posteriors <- rstan::extract(ensembleFit)

log_true_ssbSamples <- posteriors$log_true_ssb
true_ssbSamples <- exp(log_true_ssbSamples)

# Calculate posterior summaries for true state
true_ssb_summary <- data.frame(
  year = yrs,
  median_ssb = apply(true_ssbSamples, 2, median),
  mean_ssb = apply(true_ssbSamples, 2, mean),
  lower_95 = apply(true_ssbSamples, 2, quantile, 0.025),
  upper_95 = apply(true_ssbSamples, 2, quantile, 0.975)
)

# Extract discrepancy components
gammaSamples <- posteriors$gamma  # Long-term discrepancy
etaSamples <- posteriors$eta      # Shared short-term discrepancy
zSamples <- posteriors$z          # Individual short-term discrepancy

# Summarize discrepancy components
gamma_summary <- data.frame(
  model = paste0("Model_", 1:stan_data$nModels),
  M = scen$M,
  h = scen$h,
  median_gamma = apply(gammaSamples, 2, median),
  mean_gamma = apply(gammaSamples, 2, mean),
  lower_95 = apply(gammaSamples, 2, quantile, 0.025),
  upper_95 = apply(gammaSamples, 2, quantile, 0.975)
)

eta_summary <- data.frame(
  year = yrs,
  median_eta = apply(etaSamples, 2, median),
  mean_eta = apply(etaSamples, 2, mean),
  lower_95 = apply(etaSamples, 2, quantile, 0.025),
  upper_95 = apply(etaSamples, 2, quantile, 0.975)
)
```


### Plot the Ensemble Results

```{r plot_ensemble, eval=FALSE}
# Plot the ensemble SSB estimates with uncertainty
ggplot(true_ssb_summary, aes(x = year)) +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95), fill = "grey70", alpha = 0.5) +
  geom_line(aes(y = median_ssb), linewidth = 1) +
  geom_line(data = outputs,
            aes(x = year, y = ssb, group = model_id, color = as.factor(M)),
            linetype = "dashed", alpha = 0.6) +
  labs(title = "Ensemble Stock Assessment: Spawning Stock Biomass",
       subtitle = "Solid line shows ensemble estimate with 95% credible interval",
       x = "Year",
       y = "SSB",
       color = "Natural Mortality (M)") +
  theme_minimal() +
  theme(legend.position = "bottom")
```


## Extending the Framework to Other Variables

The approach can be generalized to create ensembles for other stock assessment variables beyond SSB.

```{r generalized_function}
# Function to extract any variable from stock assessment models
getSSVar <- function(M, h, variable = "SpawnBio", path = getwd()) {
  # Load the saved SS output
  load(file.path(path, M, h, "ss.RData"))
  
  # Extract time series for the requested variable
  if (variable %in% names(ss$timeseries)) {
    ts <- ss$timeseries %>%
      dplyr::select(Yr, !!sym(variable)) %>% 
      dplyr::rename(year = Yr, value = !!sym(variable))
  } else {
    stop(paste("Variable", variable, "not found in the model output timeseries"))
  }
  
  # Extract uncertainty if available
  # Determine pattern for matching in derived_quants
  pattern <- paste0("^", variable, "_")
  if (variable == "SpawnBio") pattern <- "^SSB_"  # Special case for SSB
  if (variable == "F") pattern <- "^F_" # Special case for F
  
  if ("derived_quants" %in% names(ss)) {
    sd <- ss$derived_quants %>%
      filter(grepl(pattern, Label)) %>%
      mutate(year = as.numeric(gsub(pattern, "", Label))) %>%
      dplyr::select(year, StdDev) %>%
      dplyr::rename(value_sd = StdDev)
    
    ts <- left_join(ts, sd, by = "year")
  } else {
    # If standard deviation not available, use 10% as rough estimate
    ts$value_sd <- ts$value * 0.1
  }
  
  # Add model identifiers
  ts$M <- M
  ts$h <- h
  ts$variable <- variable
  
  return(ts)
}
```


## Creating a Config-Based Ensemble Approach

```{r config_approach}
# Function to prepare data for ensemble model
prepareEnsembleData <- function(scen, variable = "SpawnBio", path = getwd()) {
  # Apply function to all models in the grid
  outputs <- list()
  for (i in 1:nrow(scen)) {
    M <- scen[i, "M"]
    h <- scen[i, "h"]
    outputs[[paste0("M", M, "_h", h)]] <- getSSVar(M, h, variable, path)
  }
  
  # Combine all outputs
  outputs <- bind_rows(outputs, .id = "model_id")
  outputs <- outputs %>%
    mutate(logValue = log(value),
           logSD = value_sd / value)  # Delta method for SD on log scale
  
  # Filter to common time period
  yrs <- outputs %>%
    group_by(year) %>%
    dplyr::summarize(count = dplyr::n()) %>%
    filter(count == nrow(scen)) %>%
    pull(year)
  
  outputs <- outputs %>% 
    filter(year %in% yrs) %>% 
    filter(!is.na(value_sd))
  
  return(list(data = outputs, years = yrs))
}

# Function to run ensemble for a specific variable
runEnsemble <- function(variable, path = getwd()) {
  # Prepare data
  prep <- prepareEnsembleData(scen, variable = variable, path = path)
  outputs <- prep$data
  yrs <- prep$years
  
  # Prepare Stan data
  stan_data <- list(
    nObs = nrow(outputs),
    nYrs = length(yrs),
    nModels = nrow(scen),
    
    yearIdx = match(outputs$year, yrs),
    modelIdx = as.numeric(factor(outputs$model_id)),
    
    logValue = outputs$logValue,
    logSD = outputs$logSD
  )
  
  # Fit the model
  modelFile <- paste0("discrepancy_model_", variable, ".stan")
  writeLines(stanCode, modelFile)
  
  ensembleFit <- stan(
    file = modelFile,
    data = stan_data,
    iter = 2000,
    warmup = 1000,
    chains = 4,
    refresh = 200,
    control = list(adapt_delta = 0.9, max_treedepth = 12)
  )
  
  # Extract results
  posteriors <- rstan::extract(ensembleFit)
  
  log_true_valueSamples <- posteriors$log_true_ssb
  true_valueSamples <- exp(log_true_valueSamples)
  
  # Calculate posterior summaries
  true_value_summary <- data.frame(
    year = yrs,
    variable = variable,
    median_value = apply(true_valueSamples, 2, median),
    mean_value = apply(true_valueSamples, 2, mean),
    lower_95 = apply(true_valueSamples, 2, quantile, 0.025),
    upper_95 = apply(true_valueSamples, 2, quantile, 0.975)
  )
  
  # Return results
  return(list(
    fit = ensembleFit,
    summary = true_value_summary,
    data = outputs,
    posteriors = posteriors
  ))
}

# Function to plot ensemble results
plotEnsembleVariable <- function(ensemble_result, title = NULL) {
  variable <- unique(ensemble_result$summary$variable)
  
  if (is.null(title)) {
    title <- paste("Ensemble", variable, "Estimate")
  }
  
  ggplot(ensemble_result$summary, aes(x = year)) +
    geom_ribbon(aes(ymin = lower_95, ymax = upper_95), fill = "grey70", alpha = 0.5) +
    geom_line(aes(y = median_value), linewidth = 1) +
    geom_line(data = ensemble_result$data,
              aes(x = year, y = value, group = model_id, color = as.factor(M)),
              linetype = "dashed", alpha = 0.6) +
    labs(title = title,
         subtitle = "Solid line shows ensemble estimate with 95% credible interval",
         x = "Year",
         y = variable,
         color = "Natural Mortality (M)") +
    theme_minimal() +
    theme(legend.position = "bottom")
}
```


## Advantages of Spence's Ensemble Approach

Spence's ensemble approach offers several advantages over traditional ensemble methods:

1. **Explicit Discrepancy Modeling**: Rather than simply averaging models or selecting the "best" one, it models the relationship between each assessment and the true (unknown) state, accounting for systematic biases.
2. **Temporal Structure**: By using AR(1) processes for discrepancy terms, the approach captures temporal patterns in model errors.
3. **Shared Discrepancy Recognition**: The approach acknowledges that models fitted to the same data will have correlated errors, preventing artificial confidence when similar models agree.
4. **Improved Uncertainty Quantification**: The Bayesian framework provides more comprehensive uncertainty estimates that account for both parameter uncertainty and structural differences.
5. **Superior Performance**: According to Spence's research, this approach outperformed the "best" weighting scheme for mixture models 99% of the time and weighted averages 48% of the time.

## Conclusion

Spence's ensemble framework provides a more robust approach to combining stock assessment models than traditional weighting schemes. By explicitly modeling the discrepancy between model outputs and the true state, this approach captures both long-term biases and short-term patterns of errors across models. The implementation in this document allows for creating ensembles across a grid of models with varying natural mortality (M) and steepness (h) parameters, providing more reliable estimates of stock status with properly characterized uncertainty.

This framework can be extended to other stock assessment variables beyond SSB, offering a comprehensive approach to quantifying uncertainty in fisheries management.

<div style="text-align: center">⁂</div>

[^1]: https://ppl-ai-file-upload.s3.amazonaws.com/web/direct-files/11254397/0af2e8c1-7240-4773-835a-bf1c1f7ceb3f/Spence_ModelWeighting_beamer.pdf

[^2]: https://ppl-ai-file-upload.s3.amazonaws.com/web/direct-files/11254397/04bf867f-521d-4231-99b9-f3a921438e69/Spence_ModelWeighting_text.pdf

[^3]: https://ppl-ai-file-upload.s3.amazonaws.com/web/direct-files/11254397/d7c8d718-6ff1-4fb3-adeb-4807aeded019/wkbsea-north-Mh.html

[^4]: https://ppl-ai-file-upload.s3.amazonaws.com/web/direct-files/11254397/c412252e-a7a1-49f7-baf2-07a3cc5ccf40/wkbsea-north-Mh.Rmd

[^5]: https://pplx-res.cloudinary.com/image/upload/v1741820911/user_uploads/ZsLcIUfYZvURvOA/image.jpg

[^6]: https://pplx-res.cloudinary.com/image/upload/v1741843720/user_uploads/NoBtdUOxKGMLzNc/image.jpg

[^7]: https://pplx-res.cloudinary.com/image/upload/v1741844637/user_uploads/dCRsCCMWwuAmTeB/image.jpg

