---
title: "WKBSEABASS North"
subtitle: "Advice"
author: 
  L.T. Kell
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Estimating $F_{MNY}$ for Atlantic Mackerel}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---


```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               cache     =TRUE, 
               cache.path="../cache/North/",
               fig.path  ="../figs/North-",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```

```{r, install, eval=FALSE}
# Install CRAN packages
pkgs_cran <- c("ggplot2", "ggpubr", "GGally", "grid", "plyr", "dplyr", 
               "reshape", "data.table", "stringr", "foreach", "doParallel")

install.packages(pkgs_cran)

# Install FLR packages from FLR repository
install.packages(c("FLCore", "FLBRP", "ggplotFL","ss3om"), 
                repos = "https://flr.r-universe.dev")

# Install ICES TAF package
install.packages("TAF", repos = "https://ices-tools-prod.r-universe.dev")

library(remotes)

# Install FLCandy from GitHub
remotes::install_github("laurieKell/FLCandy")

# Install kobe package
remotes::install_github("fishfollower/kobe")

# Install Stock Synthesis packages
remotes::install_github("r4ss/r4ss")
remotes::install_github("PIFSCstockassessments/ss3om")
remotes::install_github("jabbamodel/ss3diags", force = TRUE)
```

```{r, pkgs}
require(ggplot2)
require(ggpubr)
library(GGally)
library(ggpubr)
library(grid) 

library(plyr)
library(dplyr)
library(reshape)

library(TAF)
library(data.table)

library(stringr)
library(gam)

library(FLCore)
library(FLBRP)
library(ggplotFL)
library(FLCandy)
library(kobe)

library(r4ss)
library(ss3om)
library(ss3diags)     

library(foreach)
library(doParallel)

theme_set(theme_minimal(base_size=12))
```


```{r, path}
path="P:/rfmo/ices/wkbseabass/ss3/20250307/north"
scen=expand.grid(M=seq(0.4,1.2,0.1),
                 h=c(0.8,  0.9, 0.999))
ctrl="BassIVVII.ctl"

setwd(path)
```


## Base

```{r, base, eval=!FALSE}
base=SS_output(file.path(path,"base"),verbose=FALSE,hidewarn=FALSE,printstat=FALSE) 
save(base, file=file.path(path,"base.RData"))
```

```{r, mvln}
load(file.path(path,"base.RData"))

ss  =FLCandy:::smrySS(base)
fls =FLCandy:::tryIt(buildFLSss330(base))

source("C:/active/FLCandy/R/ss-MVLN.R")
mvln=FLCandy:::mvlnFLQ(ssmvln(base[["CoVar"]],base[["derived_quants"]],mc= 1241))
```

```{r, ssb}
ggplot(as.data.frame(ssb(fls)))+
  geom_line(aes(year,data),          col="blue",linewidth=2)+
  geom_line(aes(year,ssb),data=ss$ts,col="red")
```

**Figure `r iFig=iFig+1; iFig`** 

```{r, stock}
ss$ts$stock=ss$ts$ssb/ss$refpts$ssbmsy[1]
ggplot()+
  geom_line(aes(year,stock),data=ss$kb,col="blue",linewidth=2)+
  geom_line(aes(year,stock),data=ss$ts,col="red")
```

**Figure `r iFig=iFig+1; iFig`** 

```{r, kobe}
kobe:::trendPhaseMar2(window(mvln[["stock"]],end=2022),ylim=3, ylab=expression(SSB:B[MSY]))
```

**Figure `r iFig=iFig+1; iFig`** 

```{r, harvest}
##ssb0/ssbmsy
plot(mvln$harvest)+
  geom_line(aes(year,harvest),linewidth=2,col="blue",data=ss$kb)

kobe:::trendPhaseMar2(window(mvln[["harvest"]],end=2022),ylim=3, ylab=expression(SSB:B[MSY]))
```

**Figure `r iFig=iFig+1; iFig`** 

```{r, kobes-mvln,fig.height=8, fig.width=8}
source("C:/active/FLCandy/R/ss-MVLN.R")

mvls=mdply(scen, function(M,h){
         load(file.path(path,M,h,"ss.RData"))
         set.seed(1224) 
         tryIt(model.frame(mvlnFLQ(ssmvln(ss$CoVar,ss$derived_quants,mc=500)),drop=T))})
```


```{r, ll}
source("C:/active/FLCandy/R/ss-smry.R")

scens=unnest(mlply(scen, function(M,h) {load(file.path(path,M,h,"smry.RData"));smry}))
scens=llply(scens, function(x) cbind(scen[x$Scenario,],x)) 

ggplot(transform(scens$ll,Steepness=ac(h)))+
  geom_line(aes(M,ll,col=Steepness))+
  xlab("M Multiplier")+
  ylab("Likelihood")
  #coord_cartesian(ylim=c(2400,2450))
```

**Figure `r iFig=iFig+1; iFig`** 

```{r, kobe2}
kobePhase(subset(scens$kb,year==max(year)),ylim=c(0,2),xlim=c(0,5))+
  geom_point(aes(stock,harvest),shape=21,fill="grey",col="black",size=1.5,stroke=0.01)+
  facet_grid(M~h)+
  labs(title="Current Status")
```

**Figure `r iFig=iFig+1; iFig`** Kobe

```{r, kobe3}
kobePhase(subset(mvls,year==max(year)),ylim=c(0,2),xlim=c(0,5))+
  geom_point(aes(exp(stock),exp(harvest)),shape=21,fill="grey",col="black",size=1.5,stroke=0.01)+
  facet_grid(M~h)+
  labs(title="Current Status")
```

**Figure `r iFig=iFig+1; iFig`** Kobe

```{r, kobe4}
kobe:::kobePhaseMar2(transform(subset(mvls,year==max(year)),run=paste(M,h))[,c("stock","harvest","run")])
  # geom_point(aes(stock,harvest),shape=21,fill=run,col="black",size=0.75,stroke=0.3)+
  # labs(title="Current Status",
  # colour="M")+
  # scale_x_continuous(limits=c(0,2))+
  # scale_y_continuous(limits=c(0,2))
```

**Figure `r iFig=iFig+1; iFig`** kobe 


```{r, pf}
productionFn<-function(b,r,k,p){
  t1=b*r/p
  t3=(b/k)^p
  t1*(1-t3)}

refpts<-function(r,k,p){
  data.frame(bmsy=k*(1/(p+1))^(1/p),
             msy =(r*k/((p+1)^((p+1)/p))),
             fmsy=r/((p+1)))}

scens=unnest(mlply(scen, function(M,h) {load(file.path(path,M,h,"smry.RData"));smry}))
scens=llply(scens, function(x) cbind(scen[x$Scenario,],x))

curve=ddply(scens$pfunc, .(M,h),  with,  
            data.frame(biomass=seq(0,k,length.out=101), 
                       catch  =productionFn(seq(0.001,k,length.out=101),r,k,p)))

rfs=subset(scens$refpts,quant=="hat")
rfs=cbind(rfs,with(scens$pfunc,refpts(r=r,k=k,p=p)[,c("bmsy","fmsy")]))

#                                 harvest=harvest/median(harvest))
ggplot(curve)+
  facet_grid(h~M,scale="free")+
  geom_line(aes(biomass,catch))+
  geom_rect(aes(xmin=0,    xmax=bmsy,ymin=msy,ymax=Inf),data=rfs,fill="red",   alpha=0.2)+
  geom_rect(aes(xmin=bmsy, xmax=Inf, ymin=0,  ymax=msy),data=rfs,fill="green", alpha=0.2)+
  geom_rect(aes(xmin=0,    xmax=bmsy,ymin=0,  ymax=msy),data=rfs,fill="yellow",alpha=0.2)+
  geom_rect(aes(xmin=bmsy, xmax=Inf, ymin=msy,ymax=Inf),data=rfs,fill="yellow",alpha=0.2)+
  geom_path( aes(biomass,catch,  col=-year),data=scens$sp)+
  geom_point(aes(biomass,catch),col="blue", data=subset(scens$sp,year==max(year)))+
  geom_abline(data=rfs, aes(slope=fmsy, intercept=0), 
              linetype="dashed", color="darkgrey")+
  labs(colour="M",
       title="Production Functions with Catch v Biomass",
       x = "Biomass",
       y = "Catch")+
  theme(legend.position="none")
```

**Figure `r iFig=iFig+1; iFig`** Production Functions


```{r, eql, eval=!FALSE}
source("C:/active/FLCandy/R/generic.R")
source("C:/active/FLCandy/R/ss-eqPlot.R")

dat=mlply(with(scen,file.path(path,M,h)),
               function(x)
                     tryIt(SS_output(x,verbose=FALSE,hidewarn=FALSE,printstat=FALSE)))
eql=llply(dat, function(x) tryIt(curveSS(x,maxY=2.5))) 

eql=Map(function(x) cbind(x, scen[x$Scenario,]), FLCandy:::unnest(eql[!unlist(llply(eql,is.null))]))
save(eql,file=file.path(path,"eql.RData"))
```

```{r, pf2, fig.width=12, fig.height=8}
load(file.path(path,"eql.RData"))

ggplot(eql$refpts)+
  facet_grid(h~M, labeller = labeller(
     M=function(x) paste("M ~", x),
     h=function(x) paste("h =", x)),
     scale="free",space="free")+
  # Base layers
  geom_polygon(data=eql$triangle,aes(x, y),             fill="orange",  alpha=0.5)+
  geom_rect(aes(xmin=0,    xmax=bmsy,ymin=msy,ymax=Inf),fill="#D55E00", alpha=0.3)+
  geom_rect(aes(xmin=bmsy, xmax=Inf, ymin=0,  ymax=msy),fill="#009E73", alpha=0.3)+
  geom_rect(aes(xmin=0,    xmax=bmsy,ymin=0,  ymax=msy),fill="#F0E442", alpha=0.3)+
  geom_rect(aes(xmin=bmsy, xmax=Inf, ymin=msy,ymax=Inf),fill="#F0E442", alpha=0.3)+
  coord_cartesian(ylim =c(0,7.5e3), expand = FALSE)+
  # Lines
  geom_line(data = eql$curve, aes(ssb, yield), color="blue")+
  geom_abline(data=eql$refpts, aes(slope=msy/bmsy, intercept=0), col="grey10", linetype=3) +
  # Time series
  geom_point(data = subset(eql$tseries, year %in% c(min(year), max(year))),
             aes(ssb, yield, color = factor(year)))+
  geom_point(data = subset(eql$tseries, !(year %in% c(min(year), max(year)))),
             aes(ssb, yield), color="grey10",fill="grey90", shape=21, size=1, stroke=0.2)+
  geom_path(data=eql$tseries, aes(ssb, yield),
            arrow = arrow(length = unit(0.3, "cm"), type="closed", angle=30,ends = "last"),
  linewidth = 0.25,color="grey10")+
  # Reference points
  geom_point(aes(x=bmsy, y=msy), shape=21, fill="white", linewidth=3) +
  geom_text( aes(x=bmsy, y=msy,  label="MSY"), hjust=-0.2, vjust=-0.2) +
  # Scales
  #scale_x_continuous(labels=scales::comma,breaks=seq(0, 100000,by=20000),
  #                   expand=expansion(mult=c(0, 0.05))) +
  scale_y_continuous(labels=scales::comma) +
  scale_color_manual(values= c("blue", "red"),
                    labels= c("Start", "End"),
                    name  = "Time Period") +
  # Labels
  labs(title="Yield vs SSB",
       x    =expression("Total Biomass (t)"),
       y    =expression("Production (t)"))+
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
      strip.background = element_rect(fill = "grey95"),
      strip.text = element_text(face = "bold"),
      axis.title = element_text(face = "bold"),
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.position = "bottom",
      axis.text.x=element_text(angle=45,hjust=1, size =10,margin=margin(t=10)),
      panel.grid.major.x = element_line(color = "grey90"), 
      axis.line = element_line(color = "black"))
  #geom_abline(aes(slope=msy/ssbmsy,intercept=0),data=ss$refpts[1,])
```

**Figure `r iFig=iFig+1; iFig`** Production Functions 2

```{r, kobe5}
kb=mlply(with(scen,file.path(path,M,h)),function(x)
  suppressMessages(FLCandy:::smrySS(SS_output(x,verbose=FALSE,hidewarn=FALSE,printstat=FALSE))))
kb=Map(function(x) cbind(x, scen[x$Scenario,]), unnest(kb))

kobePhase(kb$kb,xlim=c(0,5),ylim=c(0,1.25))+
    facet_grid(h~M, labeller = labeller(
          M=function(x) paste("M =", x),
          s=function(x) paste("s =", x)))+
    geom_path(aes(stock, harvest), color="blue") +
    #geom_path(data=dat$kb, aes(stock, harvest),
    #          arrow = arrow(length = unit(0.3, "cm"), type="closed", angle=30, ends = "last"),
    #          size = 0.25, color="grey10")+
    geom_point(data = subset(kb$kb, !(year %in% c(min(year), max(year)))),
          aes(stock, harvest), color="grey10", fill="grey90",
          shape=21, size=1, stroke=0.2)+
    geom_point(data = subset(kb$kb, year %in% c(min(year), max(year))),
          aes(stock, harvest, color = factor(year)))+
    scale_color_manual(values = c("blue", "red"),
      labels= c("Start", "End"), name = "Time Period")

```

### Data

#### CPUE

```{r, CPUE}
scale<-function(x,y,...){
  args=list(...)
  
  if (length(args)==0) group=rep(1,length(x)) else group=args[[1]]  
  
  gm=gam(y~lo(x)+group,data=data.frame(x=x,y=y,group=group))
  
  res=data.frame(hat =predict(gm),
                 y     =gm$y,
                 x     =x,
                 group =group,
                 scl   =c(0,coefficients(gm)[-(1:2)])[as.numeric(as.factor(group))]
  )
  res$y  =res$y  -res$scl
  res$hat=res$hat-res$scl
  
  if (length(args)==1) names(res)[4]=names(args)[1]  
  
  names(res)[1:4]=c("hat","data","year","name")
  res[,c(4,3,2,1)]}

plotIndexFn<-function(object,...){
  
  dat=FLQuants(llply(object,diags:::stdz)) 
  dat=subset(as.data.frame(dat),!is.na(data))
  rng=range(dat$year)
  dat=with(dat,scale(year,data,name=qname))
  
  ggplot(dat)+
    geom_line( aes(year,hat),data=dat[,c("year","hat")],col="grey60")+
    geom_line( aes(year,hat))+
    geom_line( aes(year,data))+
    geom_point(aes(year,data),shape=21,fill="grey")+
    facet_wrap(~name)
}

stdz<-function(x,na.rm=TRUE,mean=0) 
  if (mean==0) (x-mean(x,na.rm=na.rm))/sd(x,na.rm=na.rm) else         
               (x/mean(x,na.rm=na.rm))/sd(x,na.rm=na.rm)

u=cpueSS(base)
u=subset(ldply(u, as.data.frame, drop=T),!is.na(data))
u=ddply(u,.(.id), with, data.frame(year=year,data=data,obsStd=(data-mean(data))/sd(data)))
u=with(subset(u,!is.na(obsStd)),scale(x=year,y=obsStd,name=.id))
ggplot(u)+
  geom_line( aes(year,hat),col="grey25")+
  geom_line( aes(year,data))+
  geom_point(aes(year,data))+
  facet_grid(name~.,scale="free",space="free_x")+
  theme_bw(20)+theme(legend.position="bottom")+
  xlab("Year")+ylab("CPUE") 
```

**Figure `r iFig=iFig+1; iFig`.** Time series of CPUE indices, the continuous black line is a lowess smother showing the average trend by area; i.e. fitted to year with series as a factor. Clusters as chosen at the data prep meeting.



```{r residuals, fig.height=10,fig.width=8}
p=ggplot(u)+
  geom_hline(aes(yintercept=0))+
  geom_point(aes(year,data-hat),position=position_dodge(width = 1),colour="blue")+
  geom_point(aes(year,data-hat),position=position_dodge(width = 1),colour="red",data=subset(u,data-hat<0))+
  geom_line(aes(year,data-hat),alpha=0.5)+
  geom_linerange(aes(year,ymin=0,ymax=data-hat),position=position_dodge(width = 1),col="blue")+
  geom_linerange(aes(year,ymin=0,ymax=data-hat),position=position_dodge(width = 1),colour="red",data=subset(u,data-hat<0))+
  theme(legend.position="bottom")+
  facet_grid(name~.,scale="free",space="free_x")+
  theme_bw(20)+
  theme(legend.position="none")+
  xlab("Year")+ylab("Residuals")
```


```{r, runtest}
library(qicharts2)  

u=cpueSS(base) 
u=subset(ldply(u, as.data.frame, drop=T),!is.na(data))
u=ddply(u,.(.id), with, data.frame(year=year,data=data,obsStd=(data-mean(data))/sd(data)))
u=subset(u,!is.na(obsStd))
u=with(u,scale(x=year,y=obsStd,name=.id))
u=transform(u,residual=data-hat)
runs=ddply(u, .(name), with, qic(year,residual,chart="i")$data)
runs=ddply(runs,.(name),with, 
               data.frame(crossings=all(n.crossings>n.crossings.min),
                          runs     =all(longest.run<longest.run.max)))
runs=transform(runs,Pass=runs&crossings)

save(u,file=file.path(path,"u.RData")) 

ggplot(u)+
  geom_hline(aes(yintercept=0))+
  geom_line( aes(year,residual),position=position_dodge(width=1),col="grey60")+
  geom_point(aes(year,residual),position=position_dodge(width=1),col="red",lwd=0.5,
             data=subset(u,residual<0))+
  geom_linerange(aes(year,ymin=0,ymax=residual),position=position_dodge(width=1),col="red",lwd=0.5,
             data=subset(u,residual<0))+
  geom_point(aes(year,residual),position=position_dodge(width=1),col="black",lwd=0.5,
             data=subset(u,residual>0))+
  geom_linerange(aes(year,ymin=0,ymax=residual),position=position_dodge(width=1),col="black",lwd=0.5,
             data=subset(u,residual>0))+
  theme(legend.position="bottom")+
  facet_grid(name~.,scale="free",space="free_x")+
  theme_bw(20)+
  theme(legend.position="bottom",strip.text.y=element_text(angle=0),
                                 strip.text.x=element_text(angle=0),
                                 axis.text.x=element_text(angle=45, hjust=1))+
  xlab("Year")+ylab("")+
  geom_rect(aes(fill=Pass), xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf, alpha=0.25,
            data=runs)+
     scale_fill_manual(values=c("red","green"))+
     scale_x_continuous(breaks=seq(1950,2010,10))
```

**Figure `r iFig=iFig+1; iFig`** Residual runs tests.


```{r cpue-correlations-1, fig.height=8,fig.width=8}
load(file.path(path,"u.RData"))

gg1=ggpairs(cast(u,year~name,value="data")[,c(2:3,5:6)], 
  upper=list(continuous=wrap("cor",size=4, hjust=0.5)),
  lower=list(continuous = wrap(mydas:::my_smooth)),
  diag=list(continuous="bar"),
    title = "")+theme_bw(20)+
  theme(legend.position ="none", 
       panel.grid.major =element_blank(), 
       axis.ticks       =element_blank(), 
       axis.text.x      =element_blank(), 
       axis.text.y      =element_blank(),
       panel.border     =element_rect(linetype = 1, colour="black", fill=NA)) 

gg1
```

**Figure `r iFig=iFig+1; iFig`.** Pairwise scatter plots to examine correlations between Indices.



```{r cpue-clusters, fig.height=8,fig.width=8}
library(corrplot)

cr=cor(subset(model.frame(mcf(cpueSS(base)) ,drop=T),!is.na(data))[,-1],use="pairwise.complete.obs")  
cr[is.na(cr)]=0
corrplot(cr,diag=F,order="hclust",addrect=4)  +          
             theme(legend.position="bottom")  
```

**Figure `r iFig=iFig+1; iFig`.** Plot of the correlation matrix for the CPUE indices, blue indicate a positive correlation  and red negative. The order of the indices and the rectangular boxes are chosen based on a hierarchical cluster analysis using a set of dissimilarities for the indices.

```{r ccf, fig.height=8,fig.width=8}
cc=FLife:::my_ccf(FLQuants(dlply(u,.(name),with,as.FLQuant(data.frame(year=year,data=data)))))

ggplot(cc)+
  geom_vline(aes(xintercept=0),linetype=2,col="blue")+
  geom_linerange(aes(x=lag,ymin=0,ymax=data))+
  facet_grid(a~b)+
  theme_bw(20)+
  ylab("ACF")+xlab("Lag")
```

**Figure `r iFig=iFig+1; iFig`** Cross correlations between indices, to identify potential lags due to year-class effects.


#### Lengths


```{r, len}
lfd=suppressMessages(lenSS(base))

suppressMessages(plotLengths(lfd[[3]]))+
  facet_grid(.~year,scale="free")+
  scale_y_continuous(limits = c(NA,NA))+
  geom_vline(aes(xintercept=data),data=as.data.frame(lmean(lfd[[3]])),col="red",stroke=1.2)
```

#### Age

```{r, age}
alk=FLPar(base$ALK)
```


## Retros

```{r}
retro(dir=path,"files","retro","..",years=-10)
```

```{r}
path="P:/rfmo/ices/wkbseabass/ss3/20250307/north/retro"
scen=expand.grid(M=seq(0.4,1.2,0.1),
                 h=c(0.8,  0.9, 0.999))
ctrl="BassIVVII.ctl"

cpFiles<-function(sourceDir, destDir) {
    # Remove trailing slashes and asterisks
    sourceDir <- gsub("[/*]+$", "", sourceDir)
    
    if (.Platform$OS.type == "windows") {
      cmd=sprintf('xcopy "%s" "%s" /S /I /Y', 
                    sourceDir,
                    destDir)} else {
      cmd=sprintf("find '%s' -type f -exec cp {} '%s' \\;",
                    sourceDir,
                    destDir)}
    
    system(cmd)}

ncore=detectCores()-4  
registerDoParallel(ncore)

# Run
mScen=foreach(i=seq(dim(scen)[1]), 
                .packages=c("TAF","r4ss","ss3om","plyr","dplyr","FLCandy")) %dopar% {
                  
    h=scen[i,"h"]                   
    M=scen[i,"M"] 
    
    # Create dirs
    if (!dir.exists(file.path(path,M)))   mkdir(file.path(path,M))
    if (!dir.exists(file.path(path,M,h))) mkdir(file.path(path,M,h))
  
    # Copy files, excluding directories
    cpFiles(file.path(path,"files/*"), file.path(path,M,h))
    
    # Modify control file
    parlines=SS_parlines(file.path(path,M,h,ctrl))
  
    sln=parlines[grepl("BH_steep",c(parlines[,"Label"])),"Linenum"]
    SS_changepars(dir=file.path(path,M,h), ctlfile=ctrl, newctlfile=ctrl,linenums=seq(sln, sln),newvals=h)
  
    ##mln=parlines[grepl("NatM",    c(parlines[,"Label"])),"Linenum"]
    ##SS_changepars(dir=file.path(path,M,h), ctlfile=ctrl, newctlfile=ctrl,linenums=seq(mln, mln),newvals=M)
  
    # Modify M vector
    cFl  =scan(file.path(path,M,h,ctrl),what=as.character(),sep="\n")
    mRow=grep("# natm",tolower(cFl))
    for (iRow in mRow){
       mVec =as.numeric(unlist(strsplit(cFl[iRow], "\\s+")))
       mVec =mVec[!is.na(mVec)]*M
       cFl[iRow]=paste(mVec,collapse=" ")}
    cat(cFl,file=file.path(path,M,h,ctrl),sep="\n")
  
    # Run SS3
    r4ss::run(file.path(path,M,h), exe="ss3", show_in_console=FALSE, skipfinished=FALSE)
  
    # get SS files
    ss=FLCandy:::tryIt(SS_output(file.path(path,M,h),verbose=FALSE,hidewarn=FALSE,printstat=FALSE))
    save(ss,file=file.path(path,M,h,"ss.RData"))
  
    # Create and save FLR objects
    fls=FLCandy:::tryIt(buildFLSss330(ss))
    save(fls,file=file.path(path,M,h,"fls.RData"))
  
    # Save summary data.frames  
    smry=FLCandy:::tryIt(FLCandy:::smrySS(file.path(path,M,h)))
    save(smry,file=file.path(path,M,h,"smry.RData"))
    
    TRUE}

stopImplicitCluster()
```

```{r}
lls=mdply(scen, function(M,h) {
  load(file.path(path,M,h,"ss.RData"))
  data.frame(component=dimnames(ss$likelihoods_used)[[1]],
             ll       =ss$likelihoods_used[,-2])})

ggplot(transform(subset(lls,!(component%in%c("InitEQ_Regime","Crash_Pen"))),
                 Steepness=ac(h),
                 component=gsub("_"," ",unique(component))))+
  geom_line(aes(M,ll,col=Steepness),linewidth=1.5)+
  facet_wrap(.~component,ncol=2,scale="free")+
  xlab("M Multiplier")+
  ylab("Likelihood")+
  theme_bw()+theme(legend.position="bottom",
                   strip.text.y.left=element_text(angle=0))
```

```{r}
load("../data/inputs/Updated_stks_n82_R0_updated202501.RData")
load("../data/spp.RData")
load("../data/info.RData")
icesdata=ICESStocks
rm(ICESStocks)

bass=icesdata[4:5]
```


```{r, ts, fig.height=8}
plot(bass[["bss.27.4bc7ad-h"]])+  
   geom_flpar(data=FLPars('F'  =benchmark(bass[["bss.27.4bc7ad-h"]])[c("fmsy",    "flim","fpa")]),x=1980)+
   geom_flpar(data=FLPars('SSB'=benchmark(bass[["bss.27.4bc7ad-h"]])[c("btrigger","blim","bpa")]),x=1980)+
   xlab("Year")+ylab("")+
   theme(legend.position = "bottom", 
         plot.margin = unit(c(0, 0, 1, 0), "lines"),
         panel.background = element_rect(fill   ="NA", colour ="NA"), 
         panel.border     = element_rect(fill   ="NA", colour ="grey50"), 
         panel.grid.major = element_line(colour ="NA"), 
         panel.grid.minor = element_line(colour ="NA"))+
    facet_grid(factor(qname,levels=c("SSB","F","Catch","Rec"))~.,scale="free")
```

**Figure MSE.`r iFig=iFig+1;iFig`** Historical stock trends with reference points.