---
title: "SMA"
subtitle: "Compare likelihood profile with the Hessian-based approximation for `SR_LN(R0)`"
author: 
  L.T. Kell
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Estimating $F_{MNY}$ for Atlantic Mackerel}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---


```{r}
setwd("C:/active/scrs-papers/ss-ll/Rmd")
```


```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               cache     =TRUE, 
               cache.path="../cache/North-Mh/",
               fig.path  ="../figs/North-Mh",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```



### **Likelihood Profile for SR_LN(R0)**

Evaluate the likelihood profile for `SR_LN(R0)` 

- Use SS3's `-profile` to fix `SR_LN(R0)` at a range of values around the MLE estimate.
- For each fixed value, re-estimate other parameters to minimize the negative log-likelihood.
- Extract likelihood for each profiled point.


### **Hessian-Based Approximation**

Use the Hessian matrix to approximate the likelihood curvature around the MLE.

- Run reference case with Hessian calculation enabled (do not use `-nohess`)
- Extract the asymptotic standard error (SE) for `SR_LN(R0)` from `covar.sso`
- Construct  quadratic approximation of the likelihood

$$
\text{Log-Likelihood} \approx \text{MLE Likelihood} - \frac{(R_0 - \hat{R}_0)^2}{2 \cdot \text{SE}^2}
$$

where $\hat{R}_0$ is the MLE[^3][^4].


### **Compare**

- Plot the likelihood profile  and the Hessian-based quadratic approximation 
    - A close match suggests the likelihood is approximately quadratic, validating Hessian-derived uncertainties.
    - Deviations indicate non-quadratic behavior (e.g., skewness), requiring profile-based confidence intervals.
### **Key Diagnostics**
If the Hessian’s curvature matches the profile’s curvature at the MLE, the standard error is reliable.
A flat profile for `SR_LN(R0)` suggests poor estimability, while a sharp minimum supports precise estimation.


# SS3 profile command for SR_LN(R0)

ss3 -profile -parmtrace profile_parms.csv -probe 0.05 -num_phases 1 -phin 0.01


**Hessian Check**:
    - Verify the Hessian is invertible (no `-nohess` flag).
    - Confirm small gradients (e.g., `<1e-3`) in `CumReport.sso`[^5].
    - Non-invertible Hessian Indicates model instability; consider fixing correlated parameters or reweighting composition data.
    - Profile-Hessian Mismatch use the likelihood profile to define confidence intervals instead of Hessian SEs.

```{r}
library(r4ss)
library(ggplot2)

ss0.8=SS_output("C:/active/scrs-papers/ss-ll/data/ICCAT_2017_Model_Run_3_SSv330.23.02/00_Model_run_files/1/1/0.8")
ss0.9=SS_output("C:/active/scrs-papers/ss-ll/data/ICCAT_2017_Model_Run_3_SSv330.23.02/00_Model_run_files/1/1/0.9")
ss1.0=SS_output("C:/active/scrs-papers/ss-ll/data/ICCAT_2017_Model_Run_3_SSv330.23.02/00_Model_run_files/1/1")
ss1.1=SS_output("C:/active/scrs-papers/ss-ll/data/ICCAT_2017_Model_Run_3_SSv330.23.02/00_Model_run_files/1/1/1.1")
ss1.2=SS_output("C:/active/scrs-papers/ss-ll/data/ICCAT_2017_Model_Run_3_SSv330.23.02/00_Model_run_files/1/1/1.2")

r0Mle=c(ss0.8$parameters$Value[model$parameters$Label=="SR_LN(R0)"],
        ss0.9$parameters$Value[model$parameters$Label=="SR_LN(R0)"],
        ss1.0$parameters$Value[model$parameters$Label=="SR_LN(R0)"],
        ss1.1$parameters$Value[model$parameters$Label=="SR_LN(R0)"],
        ss1.2$parameters$Value[model$parameters$Label=="SR_LN(R0)"])

r0SE =c(ss0.8$parameters$Parm_StDev[model$parameters$Label=="SR_LN(R0)"],
        ss0.9$parameters$Parm_StDev[model$parameters$Label=="SR_LN(R0)"],
        ss1.0$parameters$Parm_StDev[model$parameters$Label=="SR_LN(R0)"],
        ss1.1$parameters$Parm_StDev[model$parameters$Label=="SR_LN(R0)"],
        ss1.2$parameters$Parm_StDev[model$parameters$Label=="SR_LN(R0)"])

ll =c(ss0.8$likelihoods_used[1,"values"],
      ss0.9$likelihoods_used[1,"values"],
      ss1.0$likelihoods_used[1,"values"],
      ss1.1$likelihoods_used[1,"values"],
      ss1.2$likelihoods_used[1,"values"])

## ll contains the NLLs for each profile point, and r0Mle the corresponding R0 values
deltaNLL=ll= min(ll)

# Quadratic curve
r0Seq    =seq(min(r0Mle), max(r0Mle), length.out=101)
quadCurve=(r0Seq-r0Mle[which.min(ll)])^2/(2*r0SE[which.min(ll)]^2)


ggplot() +
  geom_line( data=data.frame(R0=r0Seq, DeltaNLL=quadCurve),
            aes(x=R0, y=DeltaNLL),  color="red", linewidth=1) +
  geom_point(data=data.frame(R0=r0Mle, DeltaNLL=delta_NLL),
             aes(x=R0, y=DeltaNLL), color="blue", size=2) +
  geom_vline(xintercept=r0Mle[which.min(ll)], linetype="dashed") +
  ylab("Change in negative log-likelihood") +
  xlab("SR_LN(R0)")
```


```{r}
library(r4ss)
library(ggplot2)

refDir= "../data/ICCAT_2017_Model_Run_3_SSv330.23.02/00_Model_run_files/profile" 
model=SS_output(refDir, verbose=FALSE)

# Path to your profile directory (generated via SS3 -profile)
profileDir=refDir
profile=SSgetoutput(dirvec=profileDir)

# Extract profile results
profileSmry=SSplotProfile(profile,           # Profile object
                          plot=FALSE,      # Disable auto-plot
                          print=FALSE,     # No PDF creation
                          profile.string="SR_LN(R0)")$TOTAL

# Extract Hessian-Based Estimates
# Get MLE and SE from base model
r0Mle=ss$parameters$Value[model$parameters$Label=="SR_LN(R0)"]
r0SE =ss$parameters$Parm_StDev[model$parameters$Label=="SR_LN(R0)"]

# Quadratic Approximation
r0Seq=seq(min(profile_summary$Value), 
              max(profile_summary$Value), 
              length=100)

quadratic_curve=(r0Seq-r0Mle)^2/(2*r0SE^2)

ggplot() +
  # Plot likelihood profile points
  geom_point(data=profile_summary,
             aes(x=Value, y=delta_NLL),
             size=2, color="blue") +
    geom_line(data=data.frame(Value=r0Seq, delta_NLL=quadratic_curve),
            aes(x=Value, y=delta_NLL),
            color="red", linewidth=1) +
  geom_vline(xintercept=r0Mle, linetype="dashed") +
  labs(x="SR_LN(R0)", 
       y="Change in negative log-likelihood",
       title="Likelihood Profile vs. Hessian Approximation",
       subtitle=paste("MLE =", round(r0Mle, 3), 
                       "±", round(r0SE, 3), "(SE)")) +
  theme_bw()
```


Blue points: Actual likelihood profilem Red curve: Hessian-based quadratic approximation, Vertical dashed line: MLE estimate

This visual comparison helps assess whether the Hessian adequately captures the likelihood curvature for uncertainty estimation.




- Good agreement (red line ≈ blue points) near MLE → Hessian reliable
- Large deviations → Use profile confidence intervals instead

- If `r0SE` is `NA`, check `covar.sso` exists and the Hessian was inverted
- Use `model$maximum_gradient` to confirm convergence (< 0.001)
